---
title: "Twitter COVID Tweets"
author: "Timmy Miller and Darrin Weber"
date: "5/20/2021, major update: 1/21/2023"
output: html_document
resource_files:
- .env
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rtweet)
library(tidyverse)
library(rvest)
library(lubridate)
library(RcppRoll)
library(dotenv)
```


## Purpose

The purpose of this markdown is to post the latest COVID case numbers for the seven midwestern states as part of the COVID Between the Coasts reporting in partnership with WNIN and ¿Qué Pasa, Midwest? It runs every Monday at 6:30 PM CT and grabs the latest information from [USAFacts.org](https://usafacts.org/visualizations/coronavirus-covid-19-spread-map). 

This script was created and written by Timmy Miller and was updated by Darrin Weber. Darrin Weber also published the document to UE's RStudio Connect server. 

## Script

```{r}
load_dot_env()
cbc_states <- c("IN", "KY", "MI", "OH", "IL", "WI", "MN")

# Grab Twitter data
twitter_bot <- rtweet_bot(
  api_key = Sys.getenv("TWITTER_CONSUMER_KEY"),
  api_secret = Sys.getenv("TWITTER_CONSUMER_SECRET"),
  access_token = Sys.getenv("TWITTER_ACCESS_TOKEN"),
  access_secret = Sys.getenv("TWITTER_ACCESS_SECRET"))

past_tweets <- get_timeline("@WNINCovid", n = 100, token = twitter_bot)

# Grab the date of the week of the last legitimate updated COVID numbers
last_update <- past_tweets %>% 
  filter(str_detect(text, "^((Daily)|(New))")) %>% 
  select(text) %>% 
  mutate(
    date_extract = str_extract(text,
                               "by State for ([:alpha:]+ \\d{2}, \\d{4}) \\(",
                               group = 1),
    date = mdy(date_extract)
  ) %>% 
  pull(date) %>%
  max() %>%
  floor_date(unit = "week")

# Grab COVID data
covid_html_data <- 
  read_html(
    "https://usafacts.org/visualizations/coronavirus-covid-19-spread-map"
  ) %>% 
  html_nodes('a') %>%
  html_attr('href') %>% 
  str_subset("\\.csv$")

# Extracting the data from the csv files 
locate_cases <- str_which(covid_html_data, "confirmed")[1]

cases <- read_csv(covid_html_data[locate_cases],
                  col_types = cols(
                    .default = col_character(),
                    `County Name` = col_character(),
                    State = col_character()
                  )) %>% 
  rename(county_name = `County Name`)

##### Data Cleaning #####

# Getting rid of unnecessary columns/rows and filtering to the 7 states we want
cases <- select(cases, -ends_with("FIPS")) %>% 
  filter(State %in% cbc_states)

cases <- cases %>% 
  pivot_longer(!c(county_name, State), 
               names_to = "date", 
               values_to = "cases", 
               names_transform = list(date = as_date))

# Converting cases to numeric. We imported as character because of 
# potential data entry errors that used a comma as a thousands-separator.
cases$cases <- str_remove_all(cases$cases, "[:punct:]")
cases$cases <- as.numeric(cases$cases)

# Calculating the number of new cases and renaming variables 
final_covid <- cases %>% 
  rename(state = State) %>% 
  group_by(county_name, state) %>% 
  mutate(new_cases = if_else(diff(c(0, cases)) < 0, 0, diff(c(0, cases))))

### Tweet info
# We'll calculate a 3-week moving average since we're reporting every week now
# instead of every day.
covid_data_for_tweets <- final_covid %>% 
  group_by(state, week = floor_date(date, unit = "week")) %>% 
  summarize(state_new_cases = sum(new_cases), .groups = "drop_last") %>%
  mutate(moving_3wk_avg = roll_mean(state_new_cases,
                                    n = 3,
                                    fill = NA,
                                    align = "right")) %>% 
  ungroup()

# We'll only keep the dates that haven't been tweeted out yet.
manual_tweet_data <- covid_data_for_tweets %>% 
  filter(week > last_update) %>% 
  mutate(state_new_cases = 
           if_else(state_new_cases == 0,
                   "Not Reported",
                   as.character(state_new_cases)))

display_date <- stamp_date("January 22, 2020")
no_update_tweet <- 
  "No updated data from usafacts.org this week. Check back next week."

# We'll loop through each of the weeks that haven't been tweeted on yet,
# but that we have data for. But before we do, we'll need to check that
# we have new data to report on.
if(nrow(manual_tweet_data) == 0){
  post_tweet(no_update_tweet, token = twitter_bot)
  print(no_update_tweet)
} else {
  for(date_week in sort(unique(manual_tweet_data$week))){
    individual_tweet_data <- manual_tweet_data %>% 
      filter(week == date_week)
    if(all(str_detect(individual_tweet_data$state_new_cases, "Not")) == TRUE) {
      post_tweet(no_update_tweet, token = twitter_bot)
      print(no_update_tweet)
    } else {
      individual_tweet_data_final <- individual_tweet_data %>% 
        mutate(state_new_cases = if_else(
          state_new_cases == "Not Reported" & moving_3wk_avg == 0,
          "No longer reporting",
          as.character(state_new_cases)
        ),
        moving_3wk_avg = if_else(
          str_detect(state_new_cases, "^No") & moving_3wk_avg == 0,
          "",
          as.character(round(moving_3wk_avg))
        ))
      weekly_cases <- str_c(individual_tweet_data_final$state_new_cases, 
                            sep = ",")
      weekly_rolling_avg <- str_c(individual_tweet_data_final$moving_3wk_avg,
                                  sep = ",")
      states <- str_c(individual_tweet_data_final$state, sep = ",")
      tweet_initial <- str_c(states, weekly_cases, sep = ": ")
      tweet_with_space <- str_c(tweet_initial,
                                collapse = "\n",
                                " (",
                                weekly_rolling_avg,
                                ")")
      intro_to_tweet <- 
        str_c("New Weekly Reported Cases by State for ",
              display_date(max(individual_tweet_data_final$week)),
              " (3-week Moving Average in Parentheses)")
      tweet_combine <- str_c(intro_to_tweet, sep = "\n\n", tweet_with_space)
      tweet_final <- str_replace_all(tweet_combine, 
                                     "reporting \\(\\)", 
                                     "reporting")
      post_tweet(tweet_final,
                 token = twitter_bot)
      print(tweet_final)
    }
  }
}
```

