---
title: 'COVID-19: A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Chicago Zip Code Analysis"
date: "9/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
chicago_covid <- read_csv("chicago_zip_positive_cases.csv",
                          col_types = cols(Zip = col_character()))
chicago_covid <- na.omit(chicago_covid)
chicago_covid <- chicago_covid %>% 
  mutate(case_rate = Cases/Population) %>% 
  mutate(pos_rate = Cases/Tested)
zip_code <- read_csv("zip_county.csv")
zip_code_chicago <- zip_code %>% filter(primary_city == "Chicago",
                                        state == "IL")
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Chicago. Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). 

COVID-19 case numbers were obtained from the Illinois Department of Public Health
(https://www.dph.illinois.gov/covid19/covid19-statistics)


**Conclusions:** 
After analyzing various factors, we have identified a strong relationship of COVID-19 case rates to essential occupations and nonessential occupations. Moderate relationships exist between half of the variables investigated including 6 figure incomes, incomes of $30,000 and less, public transportation, public and private health insurance, and the Gini Index. Individuals reporting below the poverty line and those who are not citizens of the United States have a slight correlation. Lastly, we saw little to no relationship in people over 60 and black or African American individuals.

We see in almost every instance where ZIP code 60604 is an outlier with a case rate of over 0.05. There is a considerably smaller population within the ZIP code compared to its neighboring ZIP codes. With this being said, over 80% of ZIP 60604 has been tested for COVID-19 when this is not the case for most ZIP codes.


## Analysis

# Family Income in the Past 12 Months
```{r}
chicago_income <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B19101") %>% 
                  filter(GEOID %in% zip_code_chicago$zip,
                         !variable %in% c("B19101_001"))
chicago_income <- left_join(chicago_income, variables_2018[, 1:2], by = "variable")
chicago_income$label <- as_factor(str_replace(chicago_income$label, ".*!!(.*)", "\\1"))
```

Incomes of six figures 
```{r}
chicago_income_six_figure <- chicago_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  dplyr::select(GEOID, prop_100K) %>%
  distinct(GEOID, prop_100K)

lvls <- chicago_income_six_figure %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)
```

Incomes of $30,000 and under
```{r}
chicago_income_under_30K <- chicago_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$25,000 to $29,999",
                      "$20,000 to $24,999",
                      "$15,000 to $19,999",
                      "$10,000 to $14,999",
                      "Less than $10,000")) %>% 
  mutate(prop_30K = sum(prop)) %>% 
  dplyr::select(GEOID, prop_30K) %>% 
  distinct(GEOID, prop_30K)

lvls_30K <- chicago_income_under_30K %>% 
  arrange(prop_30K) %>% 
  pull(GEOID)

chicago_income <- left_join(chicago_income, chicago_covid, by = c("GEOID" = "Zip"))
```


```{r}
ggplot(chicago_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Income Level")
```

Deeper look into six figure incomes
```{r}
chicago_income_cor <- left_join(chicago_income_six_figure, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_income_cor, aes(prop_100K, case_rate)) +
  geom_point() +
  labs(title = "Slight Negative Correlation Between Infection and Income",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that Makes >$100,000",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_income_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_income_cor_sans60604 <- chicago_income_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_income_cor_sans60604$prop_100K, chicago_income_cor_sans60604$case_rate, use = "complete.obs")
```
Removing the outlier caused in ZIP code 60604, we see a correlation of -0.59 and p-value of 0.000001693. This p-value indicates it is statistically significant. There is a moderate negative correlation between individuals who make six figures and case rate.


Going back to individuals who make 30K and under
```{r}
chicago_income_cor_30K <- left_join(chicago_income_under_30K, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_income_cor_30K, aes(prop_30K, case_rate)) +
  geom_point() +
  labs(title = "Slight Positive Correlation Between Infection and Income",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that Makes <$30,000",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_income_cor_30K, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_income_cor_30K_sans60604 <- chicago_income_cor_30K %>% 
  filter(case_rate < .05)
cor.test(chicago_income_cor_30K_sans60604$prop_30K, chicago_income_cor_30K_sans60604$case_rate, use = "complete.obs")
```
With a significant p-value of 0.001354, we see a correlation of 0.42 between incomes of $30,000 and less and case rate.


# Poverty Status in Past 12 Months by Age
```{r}
chicago_pov <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B17020") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         variable %in% c("B17020_002", "B17020_010"))
chicago_pov <- left_join(chicago_pov, variables_2018[, 1:2], by = "variable")
chicago_pov$label <- as_factor(str_replace(chicago_pov$label, ".*!!(.*)", "\\1"))
```

Analyzing below poverty level data
```{r}
chicago_below_pov_lvl <- chicago_pov %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Income in the past 12 months below poverty level")) %>% 
  mutate(prop_below_pov = sum(prop)) %>% 
  dplyr::select(GEOID, prop_below_pov) %>% 
  distinct(GEOID, prop_below_pov)

lvls_bel_pov <- chicago_below_pov_lvl %>% 
  arrange(prop_below_pov) %>% 
  pull(GEOID)
```

```{r}
ggplot(chicago_pov) +
  geom_col(aes(factor(GEOID, levels = lvls_bel_pov), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))+
labs(title = "Poverty Distribution for Each Chicago Zip Code",
       x = "Zip Code",
       y = "Proportion of Population", 
       fill = "Poverty Level")
```

Analyzing below poverty data
```{r}
chicago_pov_cor <- left_join(chicago_below_pov_lvl, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_pov_cor, aes(prop_below_pov, case_rate)) +
  geom_point() +
  labs(title = "Slight Positive Correlation Between Infection and Poverty",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population Below the Poverty Line",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_pov_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_pov_cor_sans60604 <- chicago_pov_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_pov_cor_sans60604$prop_below_pov, chicago_pov_cor_sans60604$case_rate, use = "complete.obs")
```
Our scatterplot "fans" out (more spread out) as the proportion of the population below the poverty line increases. We see another p-value which is significant at 0.01462. We have a slight positive correlation between individuals below the poverty line and case rate of 0.33 when we exclude our outlier of ZIP code 60604.


# Occupation by Class of Worker for the Civilian Employed Population 16 Years and Over
```{r}
chicago_occ <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "C24060") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         variable %in% c("C24060_002", "C24060_003", "C24060_004", "C24060_005", "C24060_006"))
chicago_occ <- left_join(chicago_occ, variables_2018[, 1:2], by = "variable")
chicago_occ$label <- as_factor(str_replace(chicago_occ$label, ".*!!(.*)", "\\1"))
```

Analyzing Essential Occupations
```{r}
chicago_occ_essential <- chicago_occ %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Natural resources, construction, and maintenance occupations",
                      "Production, transportation, and material moving occupations")) %>% 
  mutate(prop_essential = sum(prop)) %>% 
  dplyr::select(GEOID, prop_essential) %>% 
  distinct(GEOID, prop_essential)

lvls_ess <- chicago_occ_essential %>% 
  arrange(prop_essential) %>% 
  pull(GEOID)
```

Nonessential occupations analysis
```{r}
chicago_occ_nonessential <- chicago_occ %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Management, business, science, and arts occupations",
                      "Service occupations",
                      "Sales and office occupations")) %>% 
  mutate(prop_nonessential = sum(prop)) %>% 
  dplyr::select(GEOID, prop_nonessential) %>% 
  distinct(GEOID, prop_nonessential)

lvls_noness <- chicago_occ_nonessential %>% 
  arrange(prop_nonessential) %>% 
  pull(GEOID)

chicago_occ <- left_join(chicago_occ, chicago_covid, by = c("GEOID" = "Zip"))
```


```{r}
ggplot(chicago_occ) +
  geom_col(aes(factor(GEOID, levels = lvls_ess), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupation Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Occupation")
```

Back to essential occupations
```{r}
chicago_occ_cor <- left_join(chicago_occ_essential, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_occ_cor, aes(prop_essential, case_rate)) +
  geom_point() +
  labs(title = "Moderate Positive Correlation Between Infection and Essential Occupations",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that have Essential Occupations",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_occ_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_occ_cor_sans60604 <- chicago_occ_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_occ_cor_sans60604$prop_essential, chicago_occ_cor_sans60604$case_rate, use = "complete.obs")
```
We see our strongest correlation so far comparing essential workers to the number of COVID cases with a correlation value of 0.82! The essential occupations I considered were 2 categories: number one was "Natural resources, construction, and maintenance occupations", and number two was "Production, transportation, and material moving occupations." We also see an incredibly small p-value (significant).

Let's see if we find as strong of a correlation with nonessential occupations
```{r}
chicago_occ_cor_nonessential <- left_join(chicago_occ_nonessential, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_occ_cor_nonessential, aes(prop_nonessential, case_rate)) +
  geom_point() +
  labs(title = "Moderate Negative Correlation Between Infection and Nonessential Occupations",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that has Nonessential Occupations",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_occ_cor_nonessential, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_occ_cor_nonessential_sans60604 <- chicago_occ_cor_nonessential %>% 
  filter(case_rate < .05)
cor.test(chicago_occ_cor_nonessential_sans60604$prop_nonessential, chicago_occ_cor_nonessential_sans60604$case_rate, use = "complete.obs")
```
The correlation between these two variables is -0.82, meaning as the number of nonessential workers increases, the number of COVID cases goes down and vice versa. The occupations I considered to be nonessential was "Management, business, science, and arts occupations", "Service occupations", and "Sales and office occupations." We have a significant p-value.


# Age
```{r}
chicago_age <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B01001") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B01001_001", "B01001_002", "B01001_026"))
chicago_age <- left_join(chicago_age, variables_2018[, 1:2], by = "variable")
chicago_age$label <- as_factor(str_replace(chicago_age$label, ".*!!(.*)", "\\1"))
```

Analyzing ages over sixty
```{r}
chicago_over_sixty <- chicago_age %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("60 and 61 years",
                      "62 to 64 years",
                      "65 and 66 years",
                      "67 to 69 years",
                      "70 to 74 years",
                      "75 to 79 years",
                      "80 to 84 years",
                      "85 years and over")) %>% 
  mutate(prop_60 = sum(prop)) %>% 
  dplyr::select(GEOID, prop_60) %>% 
  distinct(GEOID, prop_60)

lvls_over_60 <- chicago_over_sixty %>% 
  arrange(prop_60) %>% 
  pull(GEOID)

chicago_age <- left_join(chicago_age, chicago_covid, by = c("GEOID" = "Zip"))
```

```{r}
ggplot(chicago_age) +
  geom_col(aes(factor(GEOID, levels = lvls_over_60), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Age Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Age")
```

```{r}
chicago_age_cor <- left_join(chicago_over_sixty, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_age_cor, aes(prop_60, case_rate)) +
  geom_point() +
  labs(title = "No Correlation Between Infection and Age",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population Over 60",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_age_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_age_cor_sans60604 <- chicago_age_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_age_cor_sans60604$prop_60, chicago_age_cor_sans60604$case_rate, use = "complete.obs")
```
There is a low correlation of -0.09 between individuals over the age of 60 and COVID-19 case rates. The p-value is not significant at 0.5253.

# Race
```{r}
chicago_race <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B02001") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B02001_001"))
chicago_race <- left_join(chicago_race, variables_2018[, 1:2], by = "variable")
chicago_race$label <- as_factor(str_replace(chicago_race$label, ".*!!(.*)", "\\1"))
```

Analyzing African American data
```{r}
chicago_black <- chicago_race %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Black or African American alone")) %>% 
  mutate(prop_black = sum(prop)) %>% 
  dplyr::select(GEOID, prop_black) %>% 
  distinct(GEOID, prop_black)

lvls_black <- chicago_black %>% 
  arrange(prop_black) %>% 
  pull(GEOID)

chicago_race <- left_join(chicago_race, chicago_covid, by = c("GEOID" = "Zip"))
```

```{r}
ggplot(chicago_race) +
  geom_col(aes(factor(GEOID, levels = lvls_black), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Race Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Race")
```

Look into Black and African American data more
```{r}
chicago_race_cor <- left_join(chicago_black, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_race_cor, aes(prop_black, case_rate)) +
  geom_point() +
  labs(title = "No Correlation Between Infection and Race",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that is Black or African American",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_race_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_race_cor_sans60604 <- chicago_race_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_race_cor_sans60604$prop_black, chicago_race_cor_sans60604$case_rate, use = "complete.obs")
```
There is a small correlation of 0.08 with a p-value of 0.5661. Our p-value suggests that there is nothing significant between the African American or black race and the case rate of COVID-19.

# Nativity and Citizenship Status in the United States
```{r}
chicago_citizen <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B05001") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B05001_001"))
chicago_citizen <- left_join(chicago_citizen, variables_2018[, 1:2], by = "variable")
chicago_citizen$label <- as_factor(str_replace(chicago_citizen$label, ".*!!(.*)", "\\1"))
```

Analyzing non-citizen data
```{r}
chicago_not_citizen <- chicago_citizen %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Not a U.S. citizen")) %>% 
  mutate(prop_not_citizen = sum(prop)) %>% 
  dplyr::select(GEOID, prop_not_citizen) %>% 
  distinct(GEOID, prop_not_citizen)

lvls_not_citizen <- chicago_not_citizen %>% 
  arrange(prop_not_citizen) %>% 
  pull(GEOID)

chicago_citizen <- left_join(chicago_citizen, chicago_covid, by = c("GEOID" = "Zip"))
```

```{r}
ggplot(chicago_citizen) +
  geom_col(aes(factor(GEOID, levels = lvls_not_citizen), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Citizenship Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Citizenship Status")
```

```{r}
chicago_citizen_cor <- left_join(chicago_not_citizen, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_citizen_cor, aes(prop_not_citizen, case_rate)) +
  geom_point() +
  labs(title = "Slight Positive Correlation Between Infection and Citizenship",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that is Not a Citizen",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_citizen_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_citizen_cor_sans60604 <- chicago_citizen_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_citizen_cor_sans60604$prop_not_citizen, chicago_citizen_cor_sans60604$case_rate, use = "complete.obs")
```
There is a slight positive correlation between individuals with no citienship and the case rate of COVID-19. With a significant p-value of 0.01915, we have a correlation of 0.32.

# Means of Transportation to Work
```{r}
chicago_trans <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B08301") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B08301_001"))
chicago_trans <- left_join(chicago_trans, variables_2018[, 1:2], by = "variable")
chicago_trans$label <- as_factor(str_replace(chicago_trans$label, ".*!!(.*)", "\\1"))
```

Analyzing Public Transportation Data
```{r}
chicago_public_trans <- chicago_trans %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Public transportation (excluding taxicab)",
                      "Bus or trolley bus",
                      "Streetcar or trolley car (carro publico in Puerto Rico)",
                      "Subway or elevated",
                      "Railroad")) %>% 
  mutate(prop_public = sum(prop)) %>% 
  dplyr::select(GEOID, prop_public) %>% 
  distinct(GEOID, prop_public)

lvls_pub <- chicago_public_trans %>% 
  arrange(prop_public) %>% 
  pull(GEOID)

chicago_trans <- left_join(chicago_trans, chicago_covid, by = c("GEOID" = "Zip"))
```

```{r}
ggplot(chicago_trans) +
  geom_col(aes(factor(GEOID, levels = lvls_pub), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Transportation Distribution for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Ways of Transportation")
```

Public Transportation
```{r}
chicago_trans_cor <- left_join(chicago_public_trans, chicago_covid, by = c("GEOID"="Zip"))

ggplot(chicago_trans_cor, aes(prop_public, case_rate)) +
  geom_point() +
  labs(title = "Slight Negative Correlation Between Infection and Transportation",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Proportion of Population that Uses Public Transportation",
       y = "Case Rate (per 100,000)")+
 ggrepel::geom_label_repel(data = filter(chicago_trans_cor, case_rate > .05),
                            aes(label = GEOID))
```

```{r}
chicago_trans_cor_sans60604 <- chicago_trans_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_trans_cor_sans60604$prop_public, chicago_trans_cor_sans60604$case_rate, use = "complete.obs")
```
There is a significant p-value of 0.0002472 and a slight correlation between individuals who use public transportation and COVID-19 case rates. That correlation is -0.48. Interpreting the correlation, as more people use pubic transportation and, in Chicago's case, the L, there are less cases of COVID-19. I used "Public transportation (excluding taxicab)", "Bus or trolley bus", "Streetcar or trolley car (carro publico in Puerto Rico)", "Subway or elevated", and "Railroad" all for this variable, public transportation.



### Private Health Insurance vs. Public Health Insurance

```{r}
chicago_hi_private <- get_acs(geography = "zcta",
                            table = "B27002") %>% 
  filter(GEOID %in% zip_code_chicago$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

chicago_hi_public <- get_acs(geography = "zcta",
                           table = "B27003") %>% 
  filter(GEOID %in% zip_code_chicago$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

chicago_hi_private$label <- str_remove(chicago_hi_private$label, "Estimate!!Total!!")
chicago_hi_public$label <- str_remove(chicago_hi_public$label, "Estimate!!Total!!")

chicago_hi_private <- separate(chicago_hi_private,
                             label,
                             sep = "!!",
                             into = c("Sex", "Age", "Private_HI"))
chicago_hi_public <- separate(chicago_hi_public,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Public_HI"))

chicago_hi_private$Private_HI <- if_else(chicago_hi_private$Private_HI == "No private health insurance", "No", "Yes")
chicago_hi_public$Public_HI <- if_else(chicago_hi_public$Public_HI == "No public coverage", "No", "Yes")

chicago_hi_private$Sex <- as_factor(chicago_hi_private$Sex)
chicago_hi_public$Sex <- as_factor(chicago_hi_public$Sex)
chicago_hi_private$Age <- as_factor(chicago_hi_private$Age)
chicago_hi_public$Age <- as_factor(chicago_hi_public$Age)
chicago_hi_private$Private_HI <- as_factor(chicago_hi_private$Private_HI)
chicago_hi_public$Public_HI <- as_factor(chicago_hi_public$Public_HI) 
```

#### Private Health Insurance


```{r}
chicago_hi_private_2 <- chicago_hi_private %>% group_by(GEOID, Private_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

### Graph Private HI Coverage by proportion

lvls_chicago_hi_private <- chicago_hi_private_2 %>% 
  filter(Private_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(chicago_hi_private_2) +
  geom_col(aes(factor(GEOID, levels = lvls_chicago_hi_private), count, fill = Private_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Prevalence of Private Health Insurance for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Private Health \nInsurance Coverage?")
```



```{r}
chicago_hi_private_cor <- left_join(chicago_hi_private_2, chicago_covid, by = c("GEOID"="Zip")) %>% 
  filter(Private_HI == "Yes")

chicago_hi_private_cor %>%
  ggplot(aes(prop, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(chicago_hi_private_cor, case_rate > .05),
                                          aes(label = GEOID)) +
  labs(title = "Possession of Private Health Insurance",
       x = "Proportion of Population that has Private Health Insurance",
       y = "Case Rate (per 100,000)")
```

```{r}
chicago_hi_private_sans60604 <- chicago_hi_private_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_hi_private_sans60604$prop, chicago_hi_private_sans60604$case_rate, use = "complete.obs")
```
There is a signficant p-value with a correlation of -0.63.



#### Public Health Insurance

We breakdown each ZIP code by the proportion of residents who have public health insurance.

```{r}
chicago_hi_public_2 <- chicago_hi_public %>% group_by(GEOID, Public_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)
### Graph Public HI Coverage by proportion

lvls_chicago_hi_public <- chicago_hi_public_2 %>% 
  filter(Public_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(chicago_hi_public_2) +
  geom_col(aes(factor(GEOID, levels = lvls_chicago_hi_public), count, fill = Public_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Prevalence of Public Health Insurance for Each Chicago ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Public Health \nInsurance Coverage?")
```



```{r}
chicago_hi_public_cor <- left_join(chicago_hi_public_2, chicago_covid, by = c("GEOID"="Zip"))

filter(chicago_hi_public_cor, Public_HI == "Yes") %>%
  ggplot(aes(prop, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(chicago_hi_public_cor, 
                                          case_rate > .05,
                                          Public_HI == "Yes"),
                                          aes(label = GEOID)) +
  labs(title = "Public Health Insurance Seems to have a Slight Positive Relation to Case Rate",
       x = "Proportion of Population that has Public Health Insurance",
       y = "Case Rate (per 100,000)")
```


```{r}
chicago_hi_public_cor_sans60604 <- chicago_hi_private_cor %>% 
  filter(case_rate < .05)
cor.test(chicago_hi_public_cor_sans60604$prop, chicago_hi_public_cor_sans60604$case_rate, use = "complete.obs")
```
There is a significant p-value and a moderate negative correlation between public health insurance and case rate of -0.63.


# Gini Index
```{r}
chicago_gini <- get_acs(geography = "zcta",
                         year = 2018,
                         table = "B19083") %>% 
  filter(GEOID %in% zip_code_chicago$zip)
chicago_gini <- left_join(chicago_gini, variables_2018[, 1:2], by = "variable")
chicago_gini$label <- as_factor(str_replace(chicago_gini$label, ".*!!(.*)", "\\1"))
chicago_gini <- left_join(chicago_gini, chicago_covid, by = c("GEOID" = "Zip"))

ggplot(chicago_gini, aes(estimate, case_rate)) +
  geom_point() +
  labs(title = "Slight Negative Correlation Between Gini Index and COVID",
       subtitle = "ZIP code 60604 is an outlier.",
       x = "Gini Index",
       y = "Case Rate (per 100,000)")+
  ggrepel::geom_label_repel(data = filter(chicago_gini, case_rate > .05),
                            aes(label = GEOID))

chicago_gini_sans60604 <- filter(chicago_gini, case_rate < .05)
cor.test(chicago_gini_sans60604$estimate, chicago_gini_sans60604$case_rate, use = "complete.obs")
```
Our p-value is 0.00189 making this significant. The correlation is moderately correlated in a negative way with a value of -0.41.


## Conclusions
The ZIP code 60604 is an outlier for it's COVID-19 case rate of 0.05789. This is the highest of all the ZIPs in Chicago. This area is a part of the Chicago "loop" which full of skyscrapers, entertainment, shops, and parks. We see that over 75% of the individuals living in this area make 6 figures and over 65% employed have nonessential occupations.

We found many factors that were unsurprising with the way their correlation reflected. For example, we would have expected those who make more money are less impacted by COVID-19. Strong correlation occurred between occupations, both essential and nonessential. More moderate relationships were found in health insurance type (public vs. private), incomes (higher and lower ends), and public transportation.

There were variables that we looked into an found little to no relationship with COVID-19. Those variables include age (over 60), and race (black or African American).
