---
title: 'COVID-19: A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Chicago Service Worker Analysis"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

We answer two questions in this report: who is the Chicago service worker? and what demographics explain the COVID case rate in Chicago ZIP codes. We look at ZIP code level data in order to understand what the service worker looks like in Chicago. We also examine the available COVID data on a ZIP code level and run a statistical linear model to find the demographics that explain the variation we see from the COVID case rate. These demographics do not necessarily drive the COVID case rate, but they do help explain the differences we are seeing in the case rate from ZIP code to ZIP code. 

## Data Obtained

### COVID-19

Pearl's section

### Demographics

Pearl's section and then Ethan's section

## Who is a Chicago Service Worker?

Timmy's section. (Timmy, be sure to include a simply interpretation of who the service worker is along as well as what professions we defined as being a service worker, along with the graphic)

As we performed this analysis, much of our focus was on the typical characteristics of the service worker in Chicago. We used correlation analyses, where we compared how strongly correlated one demographic value is with the proportion of service workers for each ZIP code. Correlations are measured from -1 to 1, where a correlation of -1 means that as one variable increases, the other variable decreases. A correlation of 1 means that as one variable increases, the other variable increases. The table below shows the correlations to the proportion of service workers (prop_service) for each demographic characteristic that we analyzed. Here are a couple examples of how to read this table: the correlation between the proportion of service workers (prop_service) and the proportion of people with health insurance is -.8965770. This means that as the proportion of service workers in a ZIP code increases, the proportion of people with health insurance decreases. Futher down the table, we see a correlation value of 0.76046345 for the correlation between the proportion of service workers (prop_service) and the proportion of individuals with Medicaid (prop_medicaid). This means that as the proportion of service workers in a ZIP code increases, the proportion of people with Medicaid (as opposed to other kinds of health insurance policies) increases.
```{r}
#one row correlation matrix

yyy<-cor(x=chicago_master, y=chicago_master$prop_service)
yyy<-as.data.frame(yyy, col.names=c("Variable", "Correlation Coefficient"))
orc<-yyy%>%subset(V1>=.7 |V1<= -.7)
orc<-orc%>%subset(V1!=1)
orc<-setDT(orc, keep.rownames = TRUE)
orc<-orc %>% arrange(desc(abs(V1)))
orc<-orc%>%rename(Variable=rn, Correlation.Coefficient=V1 )
print(orc)

```

## What Demographics Explain the COVID Case Rate?

Maya's section