---
title: 'COVID-19: A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Columbus ZIP Code Analysis"
date: "Analysis performed by Maya Frederick, Timmy Miller, Ethan Morlock, Pearl Muensterman, and Darrin Weber, Ph.D."
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 999, digits = 6)
library(tidycensus)
library(tidyverse)
library(leaflet)
library(sf)
library(tigris)
library(readxl)
options(tigris_use_cache = TRUE)
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_col <- zip_code %>% filter(primary_city == "Columbus", state == "OH")
geometry_zip <- read_sf("../Data/All_zips.shp", type = 6)
geometry_zip_col <- filter(geometry_zip, GEOID %in% zip_code_col$zip)
col_covid <- read_excel("../Data/columbus_covid.xlsx")
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Columbus. Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). COVID-19 case numbers were obtained from the Columbus Public Health Department ([https://public.tableau.com/views/COVID-19OutbreakSummary_15918845768300/COVID19Summaryp3?%3Adisplay_count=y&%3Aorigin=viz_share_link&%3AshowVizHome=no]). 

Throughout the document, we report the $p$-value for several statistical tests. This does nothing more than provide a measure on the amount of evidence we have to suggest statistical significance. The closer the $p$-value is to 0 the stronger the evidence that our measure (correlation most of the time) is different than 0. 

**Conclusions:**

## Analysis

### COVID

We first look at the COVID cases by ZIP codes. The city of Columbus reported their data in ranges instead of exact numbers. To be able to conduct analysis we took the average of that particular range for that ZIP code. That is why ZIP codes of the same color have the exact same case rate. Columbus does not seem to have any outliers as there is a nice spread of the COVID case rate. As you can see the more metropolitan areas have a little bit higher case rate which is what is expected. 

```{r}
geometry_zip_col$GEOID <- as.numeric(geometry_zip_col$GEOID)
col_covid_geo <- left_join(col_covid, geometry_zip_col, by = c("ZIP" = "GEOID"))
col_covid_geo <- st_as_sf(col_covid_geo, sf_column_name = "geometry")
```

```{r}
pal <- colorNumeric(palette = "viridis", domain = col_covid_geo$average)
col_covid_geo %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", col_covid_geo$ZIP,
                            "</strong><br /> Case Rate ", signif(col_covid_geo$average, digits = 4)),
              weight = 4,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(average)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~ average,
            title = "Case Rate (per 100000)",
            opacity = 1)
```

### Income
```{r}
col_income <- get_acs(geography = "zcta",
                        table = "B19101") %>% 
  filter(GEOID %in% zip_code_col$zip,
         !variable %in% c("B19101_001"))
col_income <- left_join(col_income, variables_2018[, 1:2], by = "variable")
col_income$label <- as_factor(str_replace(col_income$label, ".*!!(.*)", "\\1"))
geometry_zip_col$GEOID <- as.character(geometry_zip_col$GEOID)
col_income <- left_join(col_income, geometry_zip_col, by = "GEOID")
col_income <- st_as_sf(col_income, sf_column_name = "geometry")
```

We look at the income data in increments designated by the Census Bureau. We calculated the proportion of people making over \$100,000 to order the ZIP codes on the $x$-axis in the following graph. 

```{r}
col_income_six_figure <- col_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(GEOID, prop_100K) %>% 
  distinct(GEOID, prop_100K)

lvls <- col_income_six_figure %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)

ggplot(col_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Distribution for Each Columbus ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Income Level")
  
```

We see a nice distribution of six-figure incomes across the ZIP codes, ranging from just over 1\% in ZIP 43217 all the way to over 70\% of ZIP code 43212 making over \$100,000. 

Next, we performed an ANOVA test. The ANOVA test tells us if there is a significant difference in income within the groups of the COVID averages. When performing tha ANOVA test, we looked at the p-value and the R^2 value. The p-value tells us if there is a significant difference between the groups of ZIP codes and the R^2 value is related to the correlation. In simple terms the R^2 value is a rough measure of how much COVID can be explained by income.

**The ANOVA test for proportion of people making over 100K resulted in a low p-value of 0.00764 and a fairly high R^2 value of 0.403. This indicates that there is some evidence to say that ZIP codes that are wealthier are correlated with having a higher COVID case rate. **

### Gini Index
```{r}
col_gini <- get_acs(geography = "zcta",
                      table = "B19083")

geometry_zip_col$GEOID <- as.character(geometry_zip_col$GEOID)
col_gini <- left_join(col_gini, geometry_zip_col, by = "GEOID")

col_gini <- st_as_sf(col_gini, sf_column_name = "geometry")

col_gini <- col_gini %>% 
  filter(GEOID %in% zip_code_col$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable")

col_gini$label <- as_factor(str_replace(col_gini$label, ".*!!(.*)", "\\1"))
```

We can view a map of the Gini Index across Columbus ZIP codes: 
```{r}
pal_gini <- colorNumeric(palette = "viridis", domain = col_gini$estimate)

col_gini %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", col_gini$GEOID,
                            "</strong><br /> Gini Index: ", col_gini$estimate),
              weight = 2,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal_gini(estimate)) %>% 
  addLegend("bottomright",
            pal = pal_gini,
            values = ~ estimate,
            title = "Gini Index",
            opacity = 1)
```

**We see that no real pattern emerges between the Gini Index and COVID. This is supported by our ANOVA test. The ANOVA test gave us a p-value that was 0.2679 and an R^2 value of 0.175. This means that the Gini Index is not correlated to the case rate of COVID.**

### Insurance Coverage 

Since access to private health insurance is not universal, we looked at the relationship between having some type of private health insurance and COVID as well as similarly looking at public health insurance. It is possible to have both public and private insurance, so these metrics are not mutually exclusive.

```{r}
col_hi_private <- get_acs(geography = "zcta",
                            table = "B27002") %>% 
  filter(GEOID %in% zip_code_col$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

col_hi_public <- get_acs(geography = "zcta",
                           table = "B27003") %>% 
  filter(GEOID %in% zip_code_col$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

col_hi_private$label <- str_remove(col_hi_private$label, "Estimate!!Total!!")
col_hi_public$label <- str_remove(col_hi_public$label, "Estimate!!Total!!")

col_hi_private <- separate(col_hi_private,
                             label,
                             sep = "!!",
                             into = c("Sex", "Age", "Private_HI"))
col_hi_public <- separate(col_hi_public,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Public_HI"))

col_hi_private$Private_HI <- if_else(
  col_hi_private$Private_HI == "No private health insurance", "No", "Yes")
col_hi_public$Public_HI <- if_else(
  col_hi_public$Public_HI == "No public coverage", "No", "Yes")

col_hi_private$Sex <- as_factor(col_hi_private$Sex)
col_hi_public$Sex <- as_factor(col_hi_public$Sex)
col_hi_private$Age <- as_factor(col_hi_private$Age)
col_hi_public$Age <- as_factor(col_hi_public$Age)
col_hi_private$Private_HI <- as_factor(col_hi_private$Private_HI)
col_hi_public$Public_HI <- as_factor(col_hi_public$Public_HI) 
```

#### Private Health Insurance

We breakdown each ZIP code by the proportion of residents who have private health insurance.
```{r}
col_hi_private_2 <- col_hi_private %>% group_by(GEOID, Private_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

### Graph Private HI Coverage by proportion

lvls_col_hi_private <- col_hi_private_2 %>% 
  filter(Private_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(col_hi_private_2) +
  geom_col(aes(factor(GEOID, levels = lvls_col_hi_private), count, fill = Private_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

As we see, there is a wide range of those populations that have private health insurance, ranging from FILL IN\% of the population in ZIP to only FILLIN\% of the residents in ZIP having private health insurance. But how does this relate to COVID cases?

**After performing the ANOVA test, we see evidence to suggest that there is a correlation with private health insurance and the COVID case rate. From the ANOVA test we got a significant p-value of 0.0004773 and a high R^2 value of FILL IN! We have strong evidence that there is a close connection between the proportion of the population that has private health insurance and the case rate for COVID. In this case, as the proportion that have private health insurance increases, we see a decrease in the case rate.**



#### Public Health Insurance

We breakdown each ZIP code by the proportion of residents who have public health insurance.
```{r}
col_hi_public <- get_acs(geography = "zcta",
                           table = "B27003") %>% 
  filter(GEOID %in% zip_code_col$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

col_hi_public$label <- str_remove(col_hi_public$label, "Estimate!!Total!!")

col_hi_public <- separate(col_hi_public,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Public_HI"))

col_hi_public$Public_HI <- if_else(
  col_hi_public$Public_HI == "No public coverage", "No", "Yes")


col_hi_public$Sex <- as_factor(col_hi_public$Sex)
col_hi_public$Age <- as_factor(col_hi_public$Age)
col_hi_public$Public_HI <- as_factor(col_hi_public$Public_HI) 

col_hi_public_2 <- col_hi_public %>% group_by(GEOID, Public_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

lvls_col_hi_public <- col_hi_public_2 %>% 
  filter(Public_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(col_hi_public_2) +
  geom_col(aes(factor(GEOID, levels = lvls_col_hi_public), count, fill = Public_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
    labs(title = "Public Health Insurance Seems to have a\nPositive Relation to COVID Case Rate",
       x = "Proportion of Population that has Public Health Insurance",
       y = "Case Rate (per 100,000)")
```

We still have a descent distinction between ZIP codes as to the proportion of their populations that have public insurance (ranges from 0\% to almost FILL IN\%). We can look at how these proportions relate to COVID in the following plot.


