---
title: 'COVID-19:  A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Minneapolis Zip Code Analysis"
date: "Analysis performed by Maya Frederick, Timmy Miller, Ethan Morlock, Pearl Muensterman, and Darrin Weber, Ph.D."
output: html_document
urlcolor: blue
---
## Data Issues
48209, 48212, 48216, 48217, 48226, 48233, 48242, 48243
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
library(leaflet)
library(stringr)
library(sf)
library(tigris)
source("../Scripts/functions.R")
options(tigris_use_cache = TRUE)
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_det <- zip_code %>% filter(primary_city == "Detroit",
                                      state == "MI")
det_covid <- read_csv("../Data/Detroit Rates.csv",
                      col_types = cols(ZIP_Code = col_character())) %>%
  filter(ZIP_Code %in% zip_code_det$zip)
det_population <- get_acs(geography = "zcta",
                           variable = "B01003_001") %>%
  filter(GEOID %in% zip_code_det$zip)
detroit_covid <- left_join(det_population, det_covid, by = c("GEOID" = "ZIP_Code"))
detroit_covid <- detroit_covid %>% 
  rename(ZIP = GEOID,
         Population = estimate)
detroit_covid <- select(detroit_covid, ZIP, Population, Deaths_per_HundThou, Cases_per_HundThou) %>%
  filter(ZIP != '48209' & ZIP != '48216' & ZIP != '48217' & ZIP != '48226' & ZIP != '48233' & ZIP != '48242' & ZIP != '48243')
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Detroit. Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). COVID-19 case numbers were obtained from the Detroit Department of Health. (https://codtableau.detroitmi.gov/t/DHD/views/CityofDetroit-PublicCOVIDDashboard/ZIPCodeCaseRatesDashboard?%3AisGuestRedirectFromVizportal=y&%3Aembed=y). 

## Summary

## Analysis

## Occupations (Essential VS. Nonessential)

```{r}
det_occupation <- get_acs(geography = "zcta",
                           table = "C24060") %>% 
  filter(GEOID %in% zip_code_det$zip,
         !variable %in% c("C24060_001"))

det_occupation <- left_join(det_occupation, variables_2018[, 1:2], by = "variable")

det_occupation$label <- as_factor(str_replace(det_occupation$label, ".*!!(.*)", "\\1"))
```

Analyzing Essential Occupations
```{r}
det_occupation_essen <- det_occupation %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Natural resources, construction, and maintenance occupations",
                      "Production, transportation, and material moving occupations")) %>% 
  mutate(prop_essen = sum(prop)) %>% 
  select(GEOID, prop_essen) %>% 
  distinct(GEOID, prop_essen)

lvls <- det_occupation_essen %>% 
  arrange(prop_essen) %>% 
  pull(GEOID)

det_occupation <- left_join(det_occupation, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_occupation) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupations for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Occupation")
```

The scatterplot show the correlation between Essential jobs and COVID cases. 
```{r}
det_occupation_essen_cor <- left_join(det_occupation_essen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_essen_cor) +
  geom_point(aes(x = prop_essen, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Essential Jobs and COVID",
       x = "Proportion of Essential Workers",
       y = "Case Rate (per 100,000 people)")
```

Test correlation between population with essential jobs to case rate of COVID. 
```{r}
cor.test(det_occupation_essen_cor$prop_essen, det_occupation_essen_cor$Cases_per_HundThou, use = "complete.obs")
```
The correlation between essential jobs and case rate is shown to be -0.297.  This negative correlation means the more essential jobs have a lower rate of contracting COVID. 

Analyzing Nonessential Occupations
```{r}
det_occupation_nonessen <- det_occupation %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Management, business, science, and arts occupations",
                      "Service occupations",
                      "Sales and office occupations")) %>% 
  mutate(prop_nonessen = sum(prop)) %>% 
  select(GEOID, prop_nonessen) %>% 
  distinct(GEOID, prop_nonessen)

lvls <- det_occupation_nonessen %>% 
  arrange(prop_nonessen) %>% 
  pull(GEOID)

det_occupation <- left_join(det_occupation, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_occupation) +
  geom_col(aes(GEOID, estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupations for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Occupation")
```

The scatterplot shows the correlation between Nonessential jobs and COVID cases for ZIP codes. 
```{r}
det_occupation_nonessen_cor <- left_join(det_occupation_nonessen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_nonessen_cor) +
  geom_point(aes(x = prop_nonessen, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Nonessential Jobs and COVID",
       x = "Proportion of Nonessential Workers",
       y = "Case Rate (per 100,000)")
```

Test correlation between population with nonessential jobs to the case rate of COVID
```{r}
cor.test(det_occupation_nonessen_cor$prop_nonessen, det_occupation_nonessen_cor$Cases_per_HundThou, use = "complete.obs")
```
The correlation between nonessential jobs and case rate is 0.297.  This means the people with nonessential jobs are more likely to get COVID. 

## Citizenship

```{r}
minn_citi <- get_acs(geography = "zcta",
                        table = "B05001") %>% 
  filter(GEOID %in% zip_code_minn$zip,
         !variable %in% c("B05001_001"))
minn_citi <- left_join(minn_citi, variables_2018[, 1:2], by = "variable")
minn_citi$label <- as_factor(str_replace(minn_citi$label, ".*!!(.*)", "\\1"))
```

Analyze citizenship
```{r}
minn_citi_status <- minn_citi %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Not a U.S. citizen")) %>% 
  mutate(prop_not_citizen = sum(prop)) %>% 
  select(GEOID, prop_not_citizen) %>% 
  distinct(GEOID, prop_not_citizen)

lvls <- minn_citi_status %>% 
  arrange(prop_not_citizen) %>% 
  pull(GEOID)

minn_citi <- left_join(minn_citi, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_citi) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Citizenship Status for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Citizenship Status")
```

The scatterplot shows the correlation between not being a U.S. citizen and COVID.
```{r}
minn_citi_cor <- left_join(minn_citi_status, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_citi_cor) +
  geom_point(aes(x = prop_not_citizen, y = case_rate)) +
  labs(title = "Correlation Between Not Being a U.S. Citizen and COVID",
       x = "Proportion of Non U.S. Citizens",
       y = "Case Rate (per 100,000)")
```

Test correlation between precentage of population that is not a U.S. citizen to case rate.
```{r}
cor.test(minn_citi_cor$prop_not_citizen, minn_citi_cor$case_rate, use = "complete.obs")
```
The correlation between being a non-citizen and case rate is 0.611.  This is a positive correlation which means a non-citizen is more likely to get COVID. 

## Income 

```{r}
minn_income <- get_acs(geography = "zcta",
                        table = "B19101") %>% 
  filter(GEOID %in% zip_code_minn$zip,
         !variable %in% c("B19101_001"))
minn_income <- left_join(minn_income, variables_2018[, 1:2], by = "variable")
minn_income$label <- as_factor(str_replace(minn_income$label, ".*!!(.*)", "\\1"))
```

Analyzing 100K and over income
```{r}
minn_income__sixfig <- minn_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(GEOID, prop_100K) %>% 
  distinct(GEOID, prop_100K)

lvls <- minn_income__sixfig %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)

minn_income <- left_join(minn_income, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Status in past year per household for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Income Status")
```

The scatterplot shows the correlation betweeen families making a six figure income or more and COVID. 
```{r}
minn_income_sixfig_cor <- left_join(minn_income__sixfig, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_income_sixfig_cor) +
  geom_point(aes(x = prop_100K, y = case_rate)) +
  labs(title = "Correlation Between Families Making 100K or More and COVID",
       x = "Proportion of Families Making 100K or More",
       y = "Case Rate (per 100,000)")
```

Test correlation between percent of population making six figure incomes to case rate of COVID
```{r}
cor.test(minn_income_sixfig_cor$prop_100K, minn_income_sixfig_cor$case_rate, use = "complete.obs")
```
The correlation between a six figure income and case rate is a correlation of -0.761.  This is a big negative correlation which means people who are making a six figure income are less likely to get COVID. 

Analyzing 30K and under income
```{r}
minn_income_30K_below <- minn_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$25,000 to $29,999",
                      "$20,000 to $24,999",
                      "$15,000 to $19,999",
                      "$10,000 to $14,999",
                      "Less than $10,000")) %>% 
  mutate(prop_30K = sum(prop)) %>% 
  select(GEOID, prop_30K) %>% 
  distinct(GEOID, prop_30K)

lvls <- minn_income_30K_below %>% 
  arrange(prop_30K) %>% 
  pull(GEOID)

minn_income <- left_join(minn_income, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))+
labs(title = "Income Status in past year per household for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Income Status")
```

The scatterplot shows the correlation between families making less than 30K and COVID. 
```{r}
minn_income_30K_below_cor <- left_join(minn_income_30K_below, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_income_30K_below_cor) +
  geom_point(aes(x = prop_30K, y = case_rate)) +
  labs(title = "Correlation Between Families Making 30K or Less and COVID",
       x = "Proportion of Families Making 30K or Less",
       y = "Case Rate (per 100,000)")
```

Test correlation between percent of population making 30K and under incomes to case rate of COVID
```{r}
cor.test(minn_income_30K_below_cor$prop_30K, minn_income_30K_below_cor$case_rate, use = "complete.obs")
```
The correlation between a smaller income and case rate is 0.702.  This is a big positive correlation.  This means poorer people who are making less money are way more likely to get COVID than people who are making a lot of money. 

## Race

```{r}
minn_race <- get_acs(geography = "zcta",
                        table = "B02001") %>% 
  filter(GEOID %in% zip_code_minn$zip,
         !variable %in% c("B02001_001"))
minn_race <- left_join(minn_race, variables_2018[, 1:2], by = "variable")
minn_race$label <- as_factor(str_replace(minn_race$label, ".*!!(.*)", "\\1"))
```

Analyzing Afican Americans in Minneapolis
```{r}
minn_AA <- minn_race %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Black or African American alone")) %>% 
  mutate(prop_AA = sum(prop)) %>% 
  select(GEOID, prop_AA) %>% 
  distinct(GEOID, prop_AA)

lvls <- minn_AA %>% 
  arrange(prop_AA) %>% 
  pull(GEOID)

minn_race <- left_join(minn_race, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_race) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Racial Status for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Racial Status")

```

The scatterplot shows the correlation between being an African American and COVID. 
```{r}
minn_race_cor <- left_join(minn_AA, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_race_cor) +
  geom_point(aes(x = prop_AA, y = case_rate)) +
  labs(title = "Correlation Between Being an African American and COVID",
       x = "Proportion of African Americans",
       y = "Case Rate (per 100,000)")
```

Test correlation between percentage of population that is African American and case rate of COVID
```{r}
cor.test(minn_race_cor$prop_AA, minn_race_cor$case_rate, use = "complete.obs")
```
The correlation between being an African American and case rate is 0.784.  This is a big positive correlation meaning African Americans are more likely to get COVID.  

## Gini Index

```{r}
minn_gini <- get_acs(geography = "zcta",
                        table = "B19083") %>% 
  filter(GEOID %in% zip_code_minn$zip)
minn_gini <- left_join(minn_gini, variables_2018[, 1:2], by = "variable")
minn_gini$label <- as_factor(str_replace(minn_gini$label, ".*!!(.*)", "\\1"))
```

Analyze Gini Index for Minneapolis.  The scatterplot shows the correlation between Gini Index and COVID.
```{r}
minn_gini <- left_join(minn_gini, minneapolis_covid, by = c("GEOID" = "ZIP"))
minn_gini <- filter(minn_gini, GEOID != '55450')

ggplot(minn_gini) +
  geom_point(aes(x = estimate, y = case_rate)) +
  labs(title = "Gini Index for Each Minneapolis Zip Code",
       x = "Gini Index",
       y = "Case Rate")

```

Test correlation between gini index and case rate of COVID
```{r}
cor.test(minn_gini$estimate, minn_gini$case_rate, use = "complete.obs")
```
The correlation between gini index and case rate is 0.059.  This is a low positive number with a p-value close to 1, so we can't conclude anything about gini index and COVID. 

## Insurance Coverage(Public VS. Private)

```{r}
minn_insurance <- get_acs(geography = "zcta",
                           table = "B18135") %>% 
  filter(GEOID %in% zip_code_minn$zip,
         !variable %in% c("B18135_023"))
minn_insurance <- left_join(minn_insurance, variables_2018[, 1:2], by = "variable")
minn_insurance$label <- as_factor(str_replace(minn_insurance$label, ".*!!(.*)", "\\1"))
```

Analyze private health insurance coverage
```{r}
minn_insurance_private <- minn_insurance %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("With private health insurance coverage")) %>% 
  mutate(prop_priv = sum(prop)) %>% 
  select(GEOID, prop_priv) %>% 
  distinct(GEOID, prop_priv)

lvls <- minn_insurance_private %>% 
  arrange(prop_priv) %>% 
  pull(GEOID)

minn_insurance <- left_join(minn_insurance, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_insurance) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Insurance Coverage for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Insurance Coverage")
```

The scatterplot shows the correlation between private health insurance and COVID.
```{r}
minn_private_cor <- left_join(minn_insurance_private, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_private_cor) +
  geom_point(aes(x = prop_priv, y = case_rate)) +
  labs(title = "Correlation Between Private Health Insurance and COVID",
       x = "Proportion of People with Private Health Insurance",
       y = "Case Rate (per 100,000)")
```

Test correlation between percentage of population with private health insurance and case rate of COVID
```{r}
cor.test(minn_private_cor$prop_priv, minn_private_cor$case_rate, use = "complete.obs")
```
The correlation between private health insurance and case rate is -0.803.  This is a big negative correlation. People with private health insurance have been contracting less cases of COVID. 

Analyze public health insurance coverage
```{r}
minn_insurance_public <- minn_insurance %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("With public health coverage")) %>% 
  mutate(prop_public = sum(prop)) %>% 
  select(GEOID, prop_public) %>% 
  distinct(GEOID, prop_public)

lvls <- minn_insurance_public %>% 
  arrange(prop_public) %>% 
  pull(GEOID)

minn_insurance <- left_join(minn_insurance, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_insurance) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Insurance Coverage for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Insurance Coverage")
```

The scatterplot shows the correlation between public health insurance and COVID. 
```{r}
minn_public_cor <- left_join(minn_insurance_public, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_public_cor) +
  geom_point(aes(x = prop_public, y = case_rate)) +
  labs(title = "Correlation Between Public Health Insurance and COVID",
       x = "Proportion of People with Public Health Insurance",
       y = "Case Rate (per 100,000)")
```

Test correlation between percentage of population with public health insurance and case rate of COVID
```{r}
cor.test(minn_public_cor$prop_public, minn_public_cor$case_rate, use = "complete.obs")
```
The correlation between public health insurance and case rate is 0.686.  This is a positive realtionship.  The people with public insurance have been contracting more cases of COVID. 

## Transportation to Work

```{r}
minn_transport <- get_acs(geography = "zcta",
                       table = "B08301") %>% 
  filter(GEOID %in% zip_code_minn$zip,
         !variable %in% c("B08301_010"))
minn_transport <- left_join(minn_transport, variables_2018[, 1:2], by = "variable")
minn_transport$label <- as_factor(str_replace(minn_transport$label, ".*!!(.*)", "\\1"))
```

Analyzing public transporation
```{r}
minn_transport_public <- minn_transport %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Public transportation (excluding taxicab)",
                      "Bus or trolley bus",
                      "Streetcar or trolley car (carro publico in Puerto Rico)",
                      "Subway or elevated",
                      "Railroad")) %>% 
  mutate(prop_pub = sum(prop)) %>% 
  select(GEOID, prop_pub) %>% 
  distinct(GEOID, prop_pub)

lvls <- minn_transport_public %>% 
  arrange(prop_pub) %>% 
  pull(GEOID)

minn_transport <- left_join(minn_transport, minneapolis_covid, by = c("GEOID" = "ZIP"))

ggplot(minn_transport) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Transportation Distribution for Each Minneapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Transportation")
```

The scatterplot shows the correlation between public transportation and COVID.
```{r}
minn_trans_cor <- left_join(minn_transport_public, minneapolis_covid, by = c("GEOID" = "ZIP"))
ggplot(minn_trans_cor) +
  geom_point(aes(x = prop_pub, y = case_rate)) +
  labs(title = "Correlation Between Public Transportation and COVID",
       x = "Proportion of People Using Public Transportation",
       y = "Case Rate (per 100,000)")
```

Test the correlation between the percentage of the population using public transportation and case rate of COVID
```{r}
minn_trans_cor <- left_join(minn_transport_public, minneapolis_covid, by = c("GEOID"="ZIP"))
cor.test(minn_trans_cor$prop_pub, minn_trans_cor$case_rate, use = "complete.obs")
```
The correlation between public transportation and case rate is 0.549.  This is a positive relationship meaning people using public transportation are more likely to get COVID.  


## Conclusions
 