---
title: 'COVID-19:  A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Detroit Zip Code Analysis"
date: "Analysis performed by Maya Frederick, Timmy Miller, Ethan Morlock, Pearl Muensterman, and Darrin Weber, Ph.D."
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
library(leaflet)
library(stringr)
library(sf)
library(tigris)
source("../Scripts/functions.R")
options(tigris_use_cache = TRUE)
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_det <- zip_code %>% filter(primary_city == "Detroit",
                                      state == "MI")
geometry_zip <- read_sf("../Data/All_zips.shp", type = 6)
geometry_zip_det <- filter(geometry_zip, GEOID %in% zip_code_det$zip)
det_covid <- read_csv("../Data/Detroit Rates.csv",
                      col_types = cols(ZIP_Code = col_character())) %>%
  filter(ZIP_Code %in% zip_code_det$zip)
det_population <- get_acs(geography = "zcta",
                           variable = "B01003_001") %>%
  filter(GEOID %in% zip_code_det$zip)
detroit_covid <- left_join(det_population, det_covid, by = c("GEOID" = "ZIP_Code"))
detroit_covid <- detroit_covid %>% 
  rename(ZIP = GEOID,
         Population = estimate)
detroit_covid <- select(detroit_covid, ZIP, Population, Deaths_per_HundThou, Cases_per_HundThou) %>%
  filter(ZIP != '48209' & ZIP != '48216' & ZIP != '48217' & ZIP != '48226' & ZIP != '48233' & ZIP != '48242' & ZIP != '48243')
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Detroit. Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). COVID-19 case numbers were obtained from the Detroit Department of Health.  This data was pulled on September 17, 2020 from these resources.  (https://codtableau.detroitmi.gov/t/DHD/views/CityofDetroit-PublicCOVIDDashboard/ZIPCodeCaseRatesDashboard?%3AisGuestRedirectFromVizportal=y&%3Aembed=y). 

## Summary

To start we filtered out the zip codes 48209, 48212, 48216, 48217, 48226, 48233, 48242, 48243.  We filtered these out because some of them had no data included with case or death rate, which is likely due to an extremely small population or a non-population zip code.  None of our demographics showed significant p-values when testing for correlations between the demographics and case rate.  This seems odd; however, populations across Detroit are all relatively equally impacted by COVID-19.  We also looked at the correaltion between death rate and the demographics: occupations, citizenship, income, and race.  The only death rates with a significant p-values were the death rates for essential and nonessential workers. The correlation between essential workers and death rate was -0.576 and the correlation between nonessential workers and death rate was 0.576.  

## Analysis

## COVID

```{r}
det_covid_geo <- left_join(detroit_covid, geometry_zip_det, by = c("ZIP" = "GEOID"))
det_covid_geo <- st_as_sf(det_covid_geo, sf_column_name = "geometry")
```

```{r}
pal <- colorNumeric(palette = "viridis", domain = det_covid_geo$Cases_per_HundThou)
det_covid_geo %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", det_covid_geo$ZIP,
                            "</strong><br /> Case Rate ", signif(det_covid_geo$Cases_per_HundThou, digits = 4)),
              weight = 4,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(Cases_per_HundThou)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~ Cases_per_HundThou,
            title = "Case Rate (per 100,000)",
            opacity = 1)
```
We see that COVID case rate is spread throughout Detroit, however ZIPs 48211, 48235, and 48207 do stand out as having a high case rate.

## Occupations (Essential VS. Nonessential)

```{r}
det_occupation <- get_acs(geography = "zcta",
                           table = "C24060") %>% 
  filter(GEOID %in% zip_code_det$zip,
         !variable %in% c("C24060_001"))

det_occupation <- left_join(det_occupation, variables_2018[, 1:2], by = "variable")

det_occupation$label <- as_factor(str_replace(det_occupation$label, ".*!!(.*)", "\\1"))
```

Many "essential" workers tend to be towards the lower end of the income bracket. We look at whether there is a relationship between essential workers and contracting COVID.

The Census Bureau breaks workers into five categories: Management, business, science and arts; service; sales and office; natural resources, construction, and maintenance; and production, transportation, and material moving. Although these categories are undoubtedly not clear distinctions between essential and non-essential, we feel a good classification is to put natural resources, construction, and maintenance and production, transportation, material moving as essential and the other three categories as non-essential. 

Analyzing Essential Occupations in Detroit.
```{r}
det_occupation_essen <- det_occupation %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Natural resources, construction, and maintenance occupations",
                      "Production, transportation, and material moving occupations")) %>% 
  mutate(prop_essen = sum(prop)) %>% 
  select(GEOID, prop_essen) %>% 
  distinct(GEOID, prop_essen)

lvls <- det_occupation_essen %>% 
  arrange(prop_essen) %>% 
  pull(GEOID)

det_occupation <- left_join(det_occupation, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_occupation) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupations for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Occupation")
```

The scatterplot shows the correlation between essential jobs and COVID cases. 
```{r}
det_occupation_essen_cor <- left_join(det_occupation_essen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_essen_cor) +
  geom_point(aes(x = prop_essen, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Essential Jobs and COVID Cases",
       x = "Proportion of Essential Workers",
       y = "Case Rate (per 100,000 people)")
```

When we test our correlation to cases, we get a value of `r cor.test(det_occupation_essen_cor$prop_essen, det_occupation_essen_cor$Cases_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_occupation_essen_cor$prop_essen, det_occupation_essen_cor$Cases_per_HundThou, use = "complete.obs")$p.value`). This tells us that there is a slightly negative correlation between the percentage of the population working "essential" jobs and the rate of COVID cases: as a ZIP code has more essential workers, it also has a lower COVID case rate.  However, the p-value is insignificant so we cannot make this assumption. 


The scatterplot shows the correlation between essential jobs and COVID deaths.
```{r}
det_occupation_essen_cor_death <- left_join(det_occupation_essen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_essen_cor_death) +
  geom_point(aes(x = prop_essen, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Essential Jobs and COVID Deaths",
       x = "Proportion of Essential Workers",
       y = "Death Rate (per 100,000 people)")
```

When we test our correlation, we get a value of `r cor.test(det_occupation_essen_cor_death$prop_essen, det_occupation_essen_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_occupation_essen_cor_death$prop_essen, det_occupation_essen_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`). This tells us that there is a moderate correlation between the percentage of the population working "essential" jobs and the rate of COVID deaths: as a ZIP code has more essential workers, it also has a lower COVID death rate.

Analyzing Nonessential Occupations in Detroit.
```{r}
det_occupation_nonessen <- det_occupation %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Management, business, science, and arts occupations",
                      "Service occupations",
                      "Sales and office occupations")) %>% 
  mutate(prop_nonessen = sum(prop)) %>% 
  select(GEOID, prop_nonessen) %>% 
  distinct(GEOID, prop_nonessen)

lvls <- det_occupation_nonessen %>% 
  arrange(prop_nonessen) %>% 
  pull(GEOID)

det_occupation <- left_join(det_occupation, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_occupation) +
  geom_col(aes(GEOID, estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupations for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Occupation")
```

The scatterplot shows the correlation between nonessential jobs and COVID cases for ZIP codes. 
```{r}
det_occupation_nonessen_cor <- left_join(det_occupation_nonessen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_nonessen_cor) +
  geom_point(aes(x = prop_nonessen, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Nonessential Jobs and COVID Cases",
       x = "Proportion of Nonessential Workers",
       y = "Case Rate (per 100,000)")
```

When we test our correlation, we get a value of `r cor.test(det_occupation_nonessen_cor$prop_nonessen, det_occupation_nonessen_cor$Cases_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_occupation_nonessen_cor$prop_nonessen, det_occupation_nonessen_cor$Cases_per_HundThou, use = "complete.obs")$p.value`). This tells us that there is a slight correlation between the percentage of the population working "nonessential" jobs and the rate of COVID cases: as a ZIP code has more nonessential workers, it also has a higher COVID case rate.  However, the p-value is insignificant so we cannot make this assumption.

The scatterplot shows the correlation between nonessential jobs and COVID deaths for ZIP codes.
```{r}
det_occupation_nonessen_cor_death <- left_join(det_occupation_nonessen, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_occupation_nonessen_cor_death) +
  geom_point(aes(x = prop_nonessen, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Nonessential Jobs and COVID Deaths",
       x = "Proportion of Nonessential Workers",
       y = "Death Rate (per 100,000)")
```

When we test our correlation, we get a value of `r cor.test(det_occupation_nonessen_cor_death$prop_nonessen, det_occupation_nonessen_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_occupation_nonessen_cor_death$prop_nonessen, det_occupation_nonessen_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`). This tells us that there is a moderate correlation between the percentage of the population working "nonessential" jobs and the rate of COVID deaths: as a ZIP code has more nonessential workers, it also has a higher COVID death rate.

## Citizenship

The Census Bureau provides a distinction between different types of U.S. citizenship (e.g., naturalization, born in U.S., born abroad, etc.). We focus on the simple distinction between being a U.S. citizen and not.

```{r}
det_citi <- get_acs(geography = "zcta",
                        table = "B05001") %>% 
  filter(GEOID %in% zip_code_det$zip,
         !variable %in% c("B05001_001"))
det_citi <- left_join(det_citi, variables_2018[, 1:2], by = "variable")
det_citi$label <- as_factor(str_replace(det_citi$label, ".*!!(.*)", "\\1"))
```

Analyzing citizenship in Detroit.
```{r}
det_citi_status <- det_citi %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Not a U.S. citizen")) %>% 
  mutate(prop_not_citizen = sum(prop)) %>% 
  select(GEOID, prop_not_citizen) %>% 
  distinct(GEOID, prop_not_citizen)

lvls <- det_citi_status %>% 
  arrange(prop_not_citizen) %>% 
  pull(GEOID)

det_citi <- left_join(det_citi, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_citi) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Citizenship Status for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Citizenship Status")
```

The scatterplot shows the correlation between not being a U.S. citizen and COVID cases in Detroit.
```{r}
det_citi_cor <- left_join(det_citi_status, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_citi_cor) +
  geom_point(aes(x = prop_not_citizen, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Not Being a U.S. Citizen in Detroit and COVID Cases",
       x = "Proportion of Non U.S. Citizens",
       y = "Case Rate (per 100,000)")
```

When we test our correlation, we get a value of `r cor.test(det_citi_cor$prop_not_citizen, det_citi_cor$Cases_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_citi_cor$prop_not_citizen, det_citi_cor$Cases_per_HundThou, use = "complete.obs")$p.value`). This is a slightly positive correlation, which means a ZIP code with a larger population of non-citizens is also showing a higher COVID case rate. However, our p-value is not significant so we really can't say this is for sure the case. 

The scatterplot shows the correlation between not being a U.S. citizen and COVID deaths in Detroit.
```{r}
det_citi_cor_death <- left_join(det_citi_status, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_citi_cor_death) +
  geom_point(aes(x = prop_not_citizen, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Not Being a U.S. Citizen in Detroit and COVID Deaths",
       x = "Proportion of Non U.S. Citizens",
       y = "Death Rate (per 100,000)")
```

When we test our correlation, we get a value of `r cor.test(det_citi_cor_death$prop_not_citizen, det_citi_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` ($p=$ `r cor.test(det_citi_cor_death$prop_not_citizen, det_citi_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`). This is a slight-to-moderate negative correlation, which means a ZIP code with a larger population of non-citizens is also showing a lower COVID death rate.  However, the p-value is insignificant so we cannot make this assumption.

## Income 

We look at the income data in increments designated by the Census Bureau. We calculated the proportion of people making over \$100,000 to order the ZIP codes on the $x$-axis in the following graph.
```{r}
det_income <- get_acs(geography = "zcta",
                        table = "B19101") %>% 
  filter(GEOID %in% zip_code_det$zip,
         !variable %in% c("B19101_001"))
det_income <- left_join(det_income, variables_2018[, 1:2], by = "variable")
det_income$label <- as_factor(str_replace(det_income$label, ".*!!(.*)", "\\1"))
```

Analyzing 100K and over income in Detroit.
```{r}
det_income__sixfig <- det_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(GEOID, prop_100K) %>% 
  distinct(GEOID, prop_100K)

lvls <- det_income__sixfig %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)

det_income <- left_join(det_income, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Status in past year per household for Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Income Status")
```

The scatterplot shows the correlation betweeen families making a six figure income or more in Detroit and COVID cases. 
```{r}
det_income_sixfig_cor <- left_join(det_income__sixfig, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_income_sixfig_cor) +
  geom_point(aes(x = prop_100K, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Families Making 100K or More in Detroit and COVID Cases",
       x = "Proportion of Families Making 100K or More",
       y = "Case Rate (per 100,000)")
```

The correlation between a six-figure income and case rate is `r cor.test(det_income_sixfig_cor$prop_100K, det_income_sixfig_cor$Cases_per_HundThou, use = "complete.obs")$estimate` ($p\approx$ `r cor.test(det_income_sixfig_cor$prop_100K, det_income_sixfig_cor$Cases_per_HundThou, use = "complete.obs")$p.value`).  This is a very slight positive correlation. ZIP codes that are wealthier are slightly correlated with having a higher COVID case rate.  However, our p-value is not significant so we cannot really make this assumption.  

The scatterplot shows the correlation between families making a six figure income or more in Detroit and COVID deaths. 
```{r}
det_income_sixfig_cor_death <- left_join(det_income__sixfig, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_income_sixfig_cor_death) +
  geom_point(aes(x = prop_100K, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Families Making 100K or More in Detroit and COVID Deaths",
       x = "Proportion of Families Making 100K or More",
       y = "Death Rate (per 100,000)")
```

The correlation between a six-figure income and death rate is `r cor.test(det_income_sixfig_cor_death$prop_100K, det_income_sixfig_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` ($p\approx$ `r cor.test(det_income_sixfig_cor_death$prop_100K, det_income_sixfig_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`).  This is a slightly positive correlation. ZIP codes that are wealthier are correlated with having a higher COVID death rate.  However, the p-value is insignificant so we cannot make this assumption. 

Analyzing 30K and under income in Detroit.
```{r}
det_income_30K_below <- det_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$25,000 to $29,999",
                      "$20,000 to $24,999",
                      "$15,000 to $19,999",
                      "$10,000 to $14,999",
                      "Less than $10,000")) %>% 
  mutate(prop_30K = sum(prop)) %>% 
  select(GEOID, prop_30K) %>% 
  distinct(GEOID, prop_30K)

lvls <- det_income_30K_below %>% 
  arrange(prop_30K) %>% 
  pull(GEOID)

det_income <- left_join(det_income, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))+
labs(title = "Income Status in past year per household for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Income Status")
```

The scatterplot shows the correlation between families making 30K or less in Detroit and COVID cases. 
```{r}
det_income_30K_below_cor <- left_join(det_income_30K_below, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_income_30K_below_cor) +
  geom_point(aes(x = prop_30K, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Families Making 30K or Less and COVID Cases",
       x = "Proportion of Families Making 30K or Less",
       y = "Case Rate (per 100,000)")
```

When we test our correlation, we get a value of (`r cor.test(det_income_30K_below_cor$prop_30K, det_income_30K_below_cor$Cases_per_HundThou, use = "complete.obs")$estimate` with $p\approx$ `r cor.test(det_income_30K_below_cor$prop_30K, det_income_30K_below_cor$Cases_per_HundThou, use = "complete.obs")$p.value`) between income and COVID.  This is a slightly negative correlation.  This means that families making 30K or less are less likely to get COVID.  However, the p-value is insignificant so we cannot make this assumption. 

The scatterplot shows the correlation between families making 30K or less in Detroit and COVID deaths.
```{r}
det_income_30K_below_cor_death <- left_join(det_income_30K_below, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_income_30K_below_cor_death) +
  geom_point(aes(x = prop_30K, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Families Making 30K or Less and COVID Deaths",
       x = "Proportion of Families Making 30K or Less",
       y = "Death Rate (per 100,000)")
```

When we test our correlation, we get a value of (`r cor.test(det_income_30K_below_cor_death$prop_30K, det_income_30K_below_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` with $p\approx$ `r cor.test(det_income_30K_below_cor_death$prop_30K, det_income_30K_below_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`) between income and COVID. This is a slightly negative correlation.  This means that families who are making 30K or less are less likely to die from COVID.  However, the p-value is insignificant so we cannot make this assumption. 

## Race

We've heard reports about minorities being hit harder by COVID. We look to see if the data supports that in Detroit.

```{r}
det_race <- get_acs(geography = "zcta",
                        table = "B02001") %>% 
  filter(GEOID %in% zip_code_det$zip,
         !variable %in% c("B02001_001"))
det_race <- left_join(det_race, variables_2018[, 1:2], by = "variable")
det_race$label <- as_factor(str_replace(det_race$label, ".*!!(.*)", "\\1"))
```

Analyzing Afican Americans in Detroit. 
```{r}
det_AA <- det_race %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Black or African American alone")) %>% 
  mutate(prop_AA = sum(prop)) %>% 
  select(GEOID, prop_AA) %>% 
  distinct(GEOID, prop_AA)

lvls <- det_AA %>% 
  arrange(prop_AA) %>% 
  pull(GEOID)

det_race <- left_join(det_race, detroit_covid, by = c("GEOID" = "ZIP"))

ggplot(det_race) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Racial Status for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Racial Status")

```

The scatterplot shows the correlation between being an African American in Detroit and COVID cases. 
```{r}
det_race_cor <- left_join(det_AA, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_race_cor) +
  geom_point(aes(x = prop_AA, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Being an African American in Detroit and COVID Cases",
       x = "Proportion of African Americans",
       y = "Case Rate (per 100,000)")
```

The correlation between the proportion of Black and African American within a ZIP code and case rate is `r cor.test(det_race_cor$prop_AA, det_race_cor$Cases_per_HundThou, use = "complete.obs")$estimate` ($p\approx$ `r cor.test(det_race_cor$prop_AA, det_race_cor$Cases_per_HundThou, use = "complete.obs")$p.value`). This is a slightly negative correlation meaning African American communities are seeing a lower case rate of COVID in Detroit.  However, the p-value is insignificant so we cannot make this assumption.

The scatterplot shows the correlation between being an African American in Detroit and COVID deaths.
```{r}
det_race_cor_death <- left_join(det_AA, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_race_cor_death) +
  geom_point(aes(x = prop_AA, y = Deaths_per_HundThou)) +
  labs(title = "Correlation Between Being an African American in Detroit and COVID Deaths",
       x = "Proportion of African Americans",
       y = "Death Rate (per 100,000)")
```

The correlation between the proportion of Black and African American within a ZIP code and death rate is `r cor.test(det_race_cor_death$prop_AA, det_race_cor_death$Deaths_per_HundThou, use = "complete.obs")$estimate` ($p\approx$ `r cor.test(det_race_cor_death$prop_AA, det_race_cor_death$Deaths_per_HundThou, use = "complete.obs")$p.value`). This is a slight positive correlation meaning African American communities are seeing a higher death rate of COVID in Detroit.  However, the p-value is insignificant so we cannot make this assumption.

## Gini Index

```{r}
det_gini <- get_acs(geography = "zcta",
                        table = "B19083") %>% 
  filter(GEOID %in% zip_code_det$zip)
det_gini <- left_join(det_gini, variables_2018[, 1:2], by = "variable")
det_gini$label <- as_factor(str_replace(det_gini$label, ".*!!(.*)", "\\1"))
```

Analyze Gini Index for Detroit  The scatterplot shows the correlation between Gini Index and COVID.
```{r}
det_gini <- left_join(det_gini, det_covid, by = c("GEOID" = "ZIP_Code"))


ggplot(det_gini) +
  geom_point(aes(x = estimate, y = Cases_per_HundThou)) +
  labs(title = "GINI Index for Each Detroit Zip Code",
       x = "Gini Index",
       y = "Case Rate")

```

Test correlation between GINI index and case rate of COVID
```{r}
gini_cor<-cor.test(det_gini$estimate, det_gini$Cases_per_HundThou, method="pearson")
gini_cor$p.value
gini_cor$estimate
```
The correlation between GINI index and case rate is -0.204, and the p value is .375, so we cannot conclude any significant relationship between these variables. 

## Insurance Coverage(Public VS. Private)
```{r}
det_hi_private <- get_acs(geography = "zcta",
                            table = "B27002") %>% 
  filter(GEOID %in% zip_code_det$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

det_hi_public <- get_acs(geography = "zcta",
                           table = "B27003") %>% 
  filter(GEOID %in% zip_code_det$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

det_hi_private$label <- str_remove(det_hi_private$label, "Estimate!!Total!!")
det_hi_public$label <- str_remove(det_hi_public$label, "Estimate!!Total!!")

det_hi_private <- separate(det_hi_private,
                             label,
                             sep = "!!",
                             into = c("Sex", "Age", "Private_HI"))
det_hi_public <- separate(det_hi_public,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Public_HI"))

det_hi_private$Private_HI <- if_else(det_hi_private$Private_HI == "No private health insurance", "No", "Yes")
det_hi_public$Public_HI <- if_else(det_hi_public$Public_HI == "No public coverage", "No", "Yes")

det_hi_private$Sex <- as_factor(det_hi_private$Sex)
det_hi_public$Sex <- as_factor(det_hi_public$Sex)
det_hi_private$Age <- as_factor(det_hi_private$Age)
det_hi_public$Age <- as_factor(det_hi_public$Age)
det_hi_private$Private_HI <- as_factor(det_hi_private$Private_HI)
det_hi_public$Public_HI <- as_factor(det_hi_public$Public_HI) 
```

```{r}
det_hi_private_2 <- det_hi_private %>% group_by(GEOID, Private_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

### Graph Private HI Coverage by proportion

lvls_det_hi_private <- det_hi_private_2 %>% 
  filter(Private_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(det_hi_private_2) +
  geom_col(aes(factor(GEOID, levels = lvls_det_hi_private), count, fill = Private_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = "bottom") +
  labs(title = "Prevalence of Private Health Insurance for Each Minneapolis ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Private Health \nInsurance Coverage?")
```

```{r}
det_hi_private_cor <- left_join(det_hi_private_2, detroit_covid, by = c("GEOID"="ZIP")) %>% 
  filter(Private_HI == "Yes")

det_hi_private_cor %>%
  ggplot(aes(prop, Cases_per_HundThou)) +
  geom_point() +
  labs(title = "Possession of Private Health Insurance Seems\nto be Strongly Related to COVID Case Rate",
       x = "Proportion of Population that has Private Health Insurance",
       y = "Case Rate (per 100,000)")
```

```{r}
private_cor<-cor.test(det_hi_private_cor$prop, det_hi_private_cor$Cases_per_HundThou, method="pearson")

private_cor$p.value
private_cor$estimate
```
The correlation is .269 and the p-value is .238, so we cannot conclude any significant relationship between private health insurance coverage and case rate.

```{r}
det_hi_public_2 <- det_hi_public %>% group_by(GEOID, Public_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)
### Graph Public HI Coverage by proportion

lvls_det_hi_public <- det_hi_public_2 %>% 
  filter(Public_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(det_hi_public_2) +
  geom_col(aes(factor(GEOID, levels = lvls_det_hi_public), count, fill = Public_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = "bottom") +
  labs(title = "Prevalence of Public Health Insurance for Each Detroit ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Public Health \nInsurance Coverage?")
```

```{r}
det_hi_public_cor <- left_join(det_hi_public_2, detroit_covid, by = c("GEOID"="ZIP")) %>% 
  filter(Public_HI == "Yes")

det_hi_public_cor %>%
  ggplot(aes(prop, Cases_per_HundThou)) +
  geom_point() +
  labs(title = "Public Health Insurance Seems to have a\nPositive Relation to COVID Case Rate",
       x = "Proportion of Population that has Public Health Insurance",
       y = "Case Rate (per 100,000)")
```

```{r}
public_cor<-cor.test(det_hi_public_cor$prop, det_hi_public_cor$Cases_per_HundThou, method="pearson")

public_cor$p.value
public_cor$estimate
```
The correlation value is -.208 and the p-value is .365, so we cannot conclude any relationship between public health insurance coverage and case rate.

## Transportation to Work

```{r}
det_transport <- get_acs(geography = "zcta",
                       table = "B08301") %>% 
  filter(GEOID %in% zip_code_det$zip,
          variable %in% c("B08301_002", str_c("B08301_0", c(10, 16:21))))
det_transport <- left_join(det_transport, variables_2018[, 1:2], by = "variable")
det_transport$label <- as_factor(str_replace(det_transport$label, ".*!!(.*)", "\\1"))
```

Analyzing public transporation
```{r}
det_transport_public <- det_transport %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Public transportation (excluding taxicab)",
                      "Taxicab")) %>% 
  mutate(prop_pub = sum(prop)) %>% 
  select(GEOID, prop_pub) %>% 
  distinct(GEOID, prop_pub)

lvls <- det_transport_public %>% 
  arrange(prop_pub) %>% 
  pull(GEOID)



ggplot(det_transport) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Transportation Distribution for Each Detroit Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Transportation")
```

The scatterplot shows the correlation between public transportation and COVID.
```{r}
det_trans_cor <- left_join(det_transport_public, detroit_covid, by = c("GEOID" = "ZIP"))
ggplot(det_trans_cor) +
  geom_point(aes(x = prop_pub, y = Cases_per_HundThou)) +
  labs(title = "Correlation Between Public Transportation and COVID",
       x = "Proportion of People Using Public Transportation",
       y = "Case Rate (per 100,000)")
```

Test the correlation between the percentage of the population using public transportation and case rate of COVID
```{r}
det_trans_cor <- left_join(det_transport_public, detroit_covid, by = c("GEOID"="ZIP"))
trans_cor<- cor.test(det_trans_cor$prop_pub, det_trans_cor$Cases_per_HundThou, method="pearson")
trans_cor$p.value
trans_cor$estimate
```
The correlation between public transportation and case rate is -.133, and the p-value is 0.564, so we cannot conclude any significant relationship between these variables.  


## Conclusions

In Detroit we looked at 7 categories and analyzed 10 variables within them by zip code.  We removed zip codes 48209, 48212, 48216, 48217, 48226, 48233, 48242, 48243 due to them having a tiny population or being a non-population zip code. 

We did not have any significant correlations between our 10 variables and case rate.  We looked at the death rate for the variables occupations, citizenship, income, and race.  The only significant correlations we had there were the correlations for essential and nonessential occupations with death rate.  This shows that Detroit may have an equal impact across all zip codes for COVID-19.  
 