---
title: 'COVID-19: A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Indianapolis Zip Code Analysis"
date: "9/4/2020"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidycensus)
library(tidyverse)
library(leaflet)
library(stringr)
library(sf)
library(tigris)
library(readxl)

options(tigris_use_cache = TRUE)
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_indy <- zip_code %>% filter(primary_city == "Indianapolis",
                                     state == "IN")
indiana_cases_by_zip <- read_excel("../Data/indiana_cases_by_zip.xlsx")
# Filtering the cases by zip to the zip codes just for Indianapolis. 
indiana_cases_by_zip <- indiana_cases_by_zip %>% filter(ZIP_CD %in% zip_code_indy$zip)

indy_covid <- read_excel("../Data/indiana_cases_by_zip.xlsx")
indy_covid <- indy_covid %>% rename(zip = ZIP_CD,
                                    population = POPULATION,
                                    cases = PATIENT_COUNT,
                                    percentage = PERCENTAGE)

indy_covid$zip <- as.character(indy_covid$zip)
indy_covid$cases <- as.numeric(indy_covid$cases)
indy_covid$population <- as.numeric(indy_covid$population)
indy_covid <- indy_covid %>% mutate(case_rate = cases/population)

indy_data_income <- get_acs(geography = "zcta",
                         year = 2018,
                         table = "B19101",
                         geometry = TRUE) %>% 
                         filter(GEOID %in% zip_code_indy$zip, !variable %in% c("B19101_001"))

indy_data_income <- left_join(indy_data_income, variables_2018[, 1:2], by = "variable")
indy_data_income$label <- as_factor(str_replace(indy_data_income$label, ".*!!(.*)", "\\1"))

indy_data_income <- indy_data_income[,-c(2,5)] # got rid of the extra zipcode and moe columns

# Making each income level it's own row then combining all the datasets
less_than_10 <- indy_data_income %>% filter(label == "Less than $10,000") 
colnames(less_than_10)[3] <- ("< $10,000")

ten_to_14 <- indy_data_income %>% filter(label == "$10,000 to $14,999") 
colnames(ten_to_14)[3] <- ("$10,000 to $14,999")

fifteen_to_20 <- indy_data_income %>% filter(label == "$15,000 to $19,999") 
colnames(fifteen_to_20)[3] <- ("$15,000 to $19,999")

twenty_to_25 <- indy_data_income %>% filter(label == "$20,000 to $24,999") 
colnames(twenty_to_25)[3] <- ("$20,000 to $24,999")

twentyfive_to_30 <- indy_data_income %>% filter(label == "$25,000 to $29,999") 
colnames(twentyfive_to_30)[3] <- ("$25,000 to $29,999")

thirty_to_35 <- indy_data_income %>% filter(label == "$30,000 to $34,999") 
colnames(thirty_to_35)[3] <- ("$30,000 to $34,999")

thirtyfive_to_40 <- indy_data_income %>% filter(label == "$35,000 to $39,999") 
colnames(thirtyfive_to_40)[3] <- ("$35,000 to $39,999")

forty_to_45 <- indy_data_income %>% filter(label == "$40,000 to $44,999") 
colnames(forty_to_45)[3] <- ("$40,000 to $44,999")

fortyfive_to_50 <- indy_data_income %>% filter(label == "$45,000 to $49,999") 
colnames(fortyfive_to_50)[3] <- ("$45,000 to $49,999")

fifty_to_60 <- indy_data_income %>% filter(label == "$50,000 to $59,999") 
colnames(fifty_to_60)[3] <- ("$50,000 to $59,999")

sixty_to_75 <- indy_data_income %>% filter(label == "$60,000 to $74,999") 
colnames(sixty_to_75)[3] <- ("$60,000 to $74,999")

seventyfive_to_100 <- indy_data_income %>% filter(label == "$75,000 to $99,999") 
colnames(seventyfive_to_100)[3] <- ("$75,000 to $99,999")

onehundred_to_125 <- indy_data_income %>% filter(label == "$100,000 to $124,999") 
colnames(onehundred_to_125)[3] <- ("$100,000 to $124,999")

onehundred25_to_150 <- indy_data_income %>% filter(label == "$125,000 to $149,999") 
colnames(onehundred25_to_150)[3] <- ("$125,000 to $149,999")

onehundred50_to_200 <- indy_data_income %>% filter(label == "$150,000 to $199,999") 
colnames(onehundred50_to_200)[3] <- ("$150,000 to $199,999")

greater_than_200 <- indy_data_income %>% filter(label == "$200,000 or more") 
colnames(greater_than_200)[3] <- ("> $200,000")

# Joining all the data sets into one by zip code
merge1 <- inner_join(less_than_10 %>% as.data.frame(), ten_to_14 %>% as.data.frame(), by = "GEOID")
merge2 <- inner_join(merge1 %>% as.data.frame(), fifteen_to_20 %>% as.data.frame(), by = "GEOID")
merge3 <- inner_join(merge2 %>% as.data.frame(), twenty_to_25 %>% as.data.frame(), by = "GEOID")
merge4 <- inner_join(merge3 %>% as.data.frame(), twentyfive_to_30 %>% as.data.frame(), by = "GEOID")
merge5 <- inner_join(merge4 %>% as.data.frame(), thirty_to_35 %>% as.data.frame(), by = "GEOID")
merge6 <- inner_join(merge5 %>% as.data.frame(), thirtyfive_to_40 %>% as.data.frame(), by = "GEOID")
merge7 <- inner_join(merge6 %>% as.data.frame(), forty_to_45 %>% as.data.frame(), by = "GEOID")
merge8 <- inner_join(merge7 %>% as.data.frame(), fortyfive_to_50 %>% as.data.frame(), by = "GEOID")
merge9 <- inner_join(merge8 %>% as.data.frame(), fifty_to_60 %>% as.data.frame(), by = "GEOID")
merge10 <- inner_join(merge9 %>% as.data.frame(), sixty_to_75 %>% as.data.frame(), by = "GEOID")
merge11 <- inner_join(merge10 %>% as.data.frame(), seventyfive_to_100 %>% as.data.frame(), by = "GEOID")
merge12 <- inner_join(merge11 %>% as.data.frame(), onehundred_to_125 %>% as.data.frame(), by = "GEOID")
merge13 <- inner_join(merge12 %>% as.data.frame(), onehundred25_to_150 %>% as.data.frame(), by = "GEOID")
merge14 <- inner_join(merge13 %>% as.data.frame(), onehundred50_to_200 %>% as.data.frame(), by = "GEOID")
## Final dataset 
indy_income_final <- inner_join(merge14 %>% as.data.frame(), greater_than_200 %>% as.data.frame(), 
                                by = "GEOID") 
# Cleaning it up 
indy_income_final <- indy_income_final[,-c(2,4,5,6,8,9,10,12,13,14,16,17,18,20,21,22,24,25,26,28,29,30,32,
                                           33,34,36,37,38,40,41,42,44,45,46,48,49,50)]
indy_income_final <- indy_income_final[,-c(15,16,17,19,20,21,23,24,25,27,28)]

## putting the income level into a different data set that has 3 brackets (low,med,high)
## determined by indystar article

indy_income_brackets <- indy_income_final %>% mutate(low = `< $10,000` + `$10,000 to $14,999` +
                                                     `$15,000 to $19,999` + `$20,000 to $24,999` +
                                                     `$25,000 to $29,999` + `$30,000 to $34,999` +
                                                     `$35,000 to $39,999`,
                                                     medium = `$40,000 to $44,999` + `$45,000 to $49,999` +
                                                     `$50,000 to $59,999` + `$60,000 to $74,999` +
                                                       `$75,000 to $99,999`,
                                                     high = `$100,000 to $124,999` + `$125,000 to $149,999`
                                                     + `$150,000 to $199,999` + `> $200,000`)
indy_income_brackets <- indy_income_brackets[,-c(2:17)]
# Combining the income brackets with the coronovirus data
colnames(indiana_cases_by_zip)[1] <- ("GEOID")
income_and_covid <- merge(indy_income_brackets, indiana_cases_by_zip, by = "GEOID")

## Want to maybe look at populate > 65% and some race demographic %'s and add those columns to the 
#  income and covid data set. 
# since they match up to be able to view everything at once. 

## race
indy_race <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B02001") %>% 
  filter(GEOID %in% zip_code_indy$zip,
         !variable %in% c("B02001_001"))
indy_race <- left_join(indy_race, variables_2018[, 1:2], by = "variable")
indy_race$label <- as_factor(str_replace(indy_race$label, ".*!!(.*)", "\\1"))

ggplot(indy_race) +
  geom_col(aes(GEOID, estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))

indy_white_race <- get_acs(geography = "zcta",
                         year = 2018,
                         table = "B02001",
                         geometry = TRUE) %>%   
                         filter(GEOID %in% zip_code_indy$zip, variable %in% c("B02001_002"))
colnames(indy_white_race)[4] <- ("white estimate")

indy_black_race <- get_acs(geography = "zcta",
                           year = 2018,
                           table = "B02001",
                           geometry = TRUE) %>%   
                           filter(GEOID %in% zip_code_indy$zip, variable %in% c("B02001_003"))
colnames(indy_black_race)[4] <- ("black estimate")

indy_asian_race <- get_acs(geography = "zcta",
                           year = 2018,
                           table = "B02001",
                           geometry = TRUE) %>%   
                           filter(GEOID %in% zip_code_indy$zip, variable %in% c("B02001_004"))
colnames(indy_asian_race)[4] <- ("asian estimate")

## merging the 3 race (black, white, asian) estimates into one data set 
merge20 <- inner_join(indy_white_race %>% as.data.frame(), 
                      indy_black_race %>% as.data.frame(), by = "GEOID")

indy_race_dem <- inner_join(merge20 %>% as.data.frame(), 
                           indy_asian_race %>% as.data.frame(), by = "GEOID")

indy_race_dem <- indy_race_dem[,-c(2,3,5,6,7,8,10,11,12,13,15,16)] # got rid of unnecesary columns

# combining the race to the indy income/covid data set 
race_income_covid <- merge(indy_race_dem, income_and_covid, by = "GEOID")

### Age demographics (Looked at 75 and up)
seventy5_to_79 <- get_acs(geography = "zcta",
                         year = 2018,
                         table = "B01001",
                         geometry = TRUE) %>%   
                         filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_023"))
colnames(seventy5_to_79)[4] <- ("75_to_79")

eighty_to_84 <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B01001",
                          geometry = TRUE) %>%   
                          filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_024"))
colnames(eighty_to_84)[4] <- ("80_to_84")

eighty5_and_up <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B01001",
                        geometry = TRUE) %>%   
                        filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_025"))
colnames(eighty5_and_up)[4] <- ("85_and_up")

# merging the above 3 datasets (they're male numbers only) and cleaning it up 
merge30 <- inner_join(seventy5_to_79 %>% as.data.frame(), eighty_to_84 %>% as.data.frame(), by = "GEOID")
male_age <- inner_join(merge30 %>% as.data.frame(), eighty5_and_up %>% as.data.frame(), by = "GEOID")

male_age <- male_age[,-c(2,3,5,6,7,8,10,11,12,13,15,16)]
male_age <- male_age %>% mutate(total_75_and_up = `75_to_79` + `80_to_84` + `85_and_up`)
male_age <- male_age[,-c(2:4)]

# Now doing female 
seventy5_to_79F <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B01001",
                          geometry = TRUE) %>%   
                          filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_047"))
colnames(seventy5_to_79F)[4] <- ("75_to_79F")

eighty_to_84F <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B01001",
                        geometry = TRUE) %>%   
                        filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_048"))
colnames(eighty_to_84F)[4] <- ("80_to_84F")

eighty5_and_upF <- get_acs(geography = "zcta",
                          year = 2018,
                          table = "B01001",
                          geometry = TRUE) %>%   
                          filter(GEOID %in% zip_code_indy$zip, variable %in% c("B01001_049"))
colnames(eighty5_and_upF)[4] <- ("85_and_upF")

# merging the above 3 datasets (they're male numbers only) and cleaning it up 
merge40 <- inner_join(seventy5_to_79F %>% as.data.frame(), eighty_to_84F %>% as.data.frame(), by = "GEOID")
female_age <- inner_join(merge40 %>% as.data.frame(), eighty5_and_upF %>% as.data.frame(), by = "GEOID")

female_age <- female_age[,-c(2,3,5,6,7,8,10,11,12,13,15,16)]
female_age <- female_age %>% mutate(total_75_and_upF = `75_to_79F` + `80_to_84F` + `85_and_upF`)
female_age <- female_age[,-c(2:4)]

# Combining male and female age (75 and up)
indy_age <- merge(male_age,female_age, by = "GEOID")
indy_age <- indy_age %>% mutate(aged_75_and_up = total_75_and_up + total_75_and_upF)
indy_age <- indy_age[,-c(2,3)]

# Combining with the race, income and covid data
age_race_income_covid <- merge(indy_age, race_income_covid, by = "GEOID")

## Converting the demographics to be proportions 
age_race_income_covid$POPULATION <- as.numeric(age_race_income_covid$POPULATION)
age_race_income_covid <- age_race_income_covid %>% mutate(`white estimate` = `white estimate`/POPULATION,
                                                          `black estimate` = `black estimate`/POPULATION,
                                                          `asian estimate` = `asian estimate`/POPULATION,
                                                          aged_75_and_up = aged_75_and_up/POPULATION,
                                                          high = high/POPULATION,
                                                          medium = medium/POPULATION,
                                                          low = low/POPULATION)


## Statistics and graphs ##
age_race_income_covid$PERCENTAGE <- as.numeric(age_race_income_covid$PERCENTAGE)
age_race_income_covid <- age_race_income_covid[-37,]  ## Got rid of the suppressed zip code
age_race_income_covid$PATIENT_COUNT <- as.numeric(age_race_income_covid$PATIENT_COUNT)
age_race_income_covid <- age_race_income_covid %>% 
  mutate(case_rate = PATIENT_COUNT/POPULATION)
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Indianapolis Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). COVID-19 case numbers were obtained from the Indiana State Department of Health. ([https://hub.mph.in.gov/dataset/covid-19-cases-by-zip])

During the process of exploratory data analysis between demographic data and COVID-19 data in each zip code in Indianapolis, we find that with income there not a correlation between having a poorer income and a higher rate of infection. Additionally, we find that when looking at race, zip codes that are majority African American have just a slightly higher case rate than zip codes that consist of majority White. Lastly, we find that there is not a correlation between having a higher age and case rate, in fact in this case, there was a correlation between the zip codes with a lower proportion of people aged 75 and up and a higher case rate. 

## Analysis

### Income
```{r, echo=FALSE}
indy_income <- get_acs(geography = "zcta",
                        table = "B19101",
                        geometry = TRUE) %>% 
  filter(GEOID %in% zip_code_indy$zip,
         !variable %in% c("B19101_001"))

indy_income <- left_join(indy_income, variables_2018[, 1:2], by = "variable")
indy_income$label <- as_factor(str_replace(indy_income$label, ".*!!(.*)", "\\1"))

indy_income_six_figure <- indy_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(GEOID, prop_100K) %>% 
  distinct(GEOID, prop_100K)

lvls <- indy_income_six_figure %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)

indy_income <- left_join(indy_income, indy_covid, by = c("GEOID" = "zip"))
```

We look at the income data in increments designated by the Census Bureau. We calculated the proportion of people making over \$100,000 to order the zip codes on the $x$-axis in the following graph. 

```{r, echo=FALSE}
ggplot(indy_income) +
  geom_col(aes(factor(GEOID, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Distribution for Each Indianapolis Zip Code",
       x = "Zip Code",
       y = "Proportion", 
       fill = "Income Level")
```

We can compare zip codes with the case rate (per 100000) of COVID in the same zip codes.
```{r,echo=FALSE}
pal <- colorNumeric(palette = "viridis", domain = indy_income$case_rate)

indy_income %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", indy_income$GEOID,
                            "</strong><br /> Case Rate ", indy_income$case_rate),
              stroke = FALSE,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(case_rate)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~ case_rate,
            title = "Case Rate (per 100000)",
            opacity = 1)
```

```{r,echo=FALSE}
indy_income_cor <- left_join(indy_income_six_figure, indy_covid, by = c("GEOID"="zip"))
```

First, we notice that there is one zip code that does not have any COVID-19 data. This zip code was suppressed. When we looked into that zip code we found that it is a popular zip code with a population of around 14,000. However, some zip codes in the COVID-19 data were suppressed, which means that it is not being reported. Having the information for that zip code could potential slightly change the results of our analysis. 
There does not seem to be any sort of correlation, i.e., the poorer the zip code the more likely to have a higher rate of infection. However, we can verify that there is no correlation by running a correlation test between the case rate and the proportion of the population that makes over \$100,000. We get a statistically non-significant ($p=$ `r cor.test(indy_income_cor$prop_100K, indy_income_cor$case_rate, use = "complete.obs")$p.value`) correlation value of `r cor.test(indy_income_cor$prop_100K, indy_income_cor$case_rate, use = "complete.obs")$estimate`. This verifies that there is not a negative correlation between case rates and higher incomes. 

Additionally, from the graph above you can see that one zip code has a much higher case rate than any of the other zip codes. That zip code happens to be 46204 and it is right in the middle of Indianapolis. This zip code has a case rate of 5.9% while the average case rate for Indianapolis zip codes is 1.85%. So 46204 has over double the average case rate for zip codes in Indianapolis. This zip code only has a population of 5,125 people which is the second to last least populous zip code. 
While not a lot of people live there, since it is right in the middle of Indianapolis there are many bars, restaurants, stores, etc., making that zip code very crowded and easy to contract the virus. Which is a possible reason why the case rate of that zip code is so high. 

## Race 

When looking at the distribution of race across Indianapolis. It can be seen that the greater majority of Indianapolis is white. 
```{r,echo=FALSE}
indy_race <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B02001") %>% 
  filter(GEOID %in% zip_code_indy$zip,
         !variable %in% c("B02001_001"))
indy_race <- left_join(indy_race, variables_2018[, 1:2], by = "variable")
indy_race$label <- as_factor(str_replace(indy_race$label, ".*!!(.*)", "\\1"))

ggplot(indy_race) +
  geom_col(aes(GEOID, estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1))
```

It can be seen that the next highest race in Indianapolis is African American. We can get the average case rate for zip codes that are majority of both races to see how they compare to each other. The average case rate for zip codes that consist of greater than 50% white people is 1.79% while the average case rate for zip codes that consists of greater than 50% African American people is 1.83%. They African American case rate is just a little bit higher than the white case rate. The overall average case rate among all zip codes is 1.85% so those averages are a little bit lower than the overall average. 

## Age
```{r,echo=FALSE}
ggplot(age_race_income_covid) + 
  geom_bar(mapping = aes(x = GEOID, y = aged_75_and_up), stat = "identity") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  ylab("Proportion of population aged 75 and up")
```

Since COVID-19 is known particularly to be more deadly towards people with comorbidity. The older population is at a much higher risk than the younger population. Which is why I focused on the population that is aged 75 and older. The average percentage of the population that is 75 and older for each zip code in Indianapolis is 5.6%. With the lowest being 2.3% and the highest being 20%. When testing for correlation between the proportion of the population aged 75 and up and the case rate, we find that there is a negative correlation. This indicates that the younger the zip code the more likely to have a higher rate of infection. 

## Conclusion
As you can see from exploring the zip code data for demographics and COVID-19 of the city of Indianapolis, we find a couple of interesting things. Lower income does not seem to be related with a higher case rate in Indianapolis. Indianapolis has a nice spread of income so it would make sense as to why income is not correlated with a higher case rate. Additionally, we can see subtle differences in case rates with regards to racial demographics. The zip codes that are majority white have just a little bit lower average case rate than the zip codes that are majority African American. Both averages are actually lower than the overall average for zip codes in Indianapolis. Lastly, when we looked into the age demographic, we focused on the population that was aged 75 or older. We found that there was actually a negative correlation between the older you are and your chances of contracting COVID-19. 