---
title: 'COVID-19: A Midwestern Reporting Project of WNIN and the University of Evansville'
author: "Louisville Zip Code Analysis"
date: "9/4/2020"
output: html_document
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 999, digits = 4)
#See louisville_eda.R for code comments
library(tidycensus)
library(tidyverse)
library(leaflet)
library(stringr)
library(sf)
library(tigris)
options(tigris_use_cache = TRUE)
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_louis <- zip_code %>% filter(primary_city == "Louisville",
                                      state == "KY")
geometry_zip <- read_sf("../Data/All_zips.shp", type = 6)
geometry_zip_louis <- filter(geometry_zip, GEOID %in% zip_code_louis$zip)
louis_covid <- read_csv("../Data/louisville_covid.csv",
                        col_types = cols(ZIP = col_character()))
louis_covid <- louis_covid %>% 
  mutate(case_rate = Cases/Population) %>% 
  rename(zip = ZIP,
         population = Population,
         cases = Cases)
```

## Background

The goal of this analysis is to find any connections, if they exist, between demographic data and the outbreak of COVID-19 in the city of Louisville. Demographic data was obtained from the Census Bureau through their 5-year American Community Survey published in 2018 (the most recent 5-year data available). COVID-19 case numbers were obtained from the Louisville Metro Department of Public Health and Wellness ([https://covid-19-in-jefferson-county-ky-lojic.hub.arcgis.com/](https://covid-19-in-jefferson-county-ky-lojic.hub.arcgis.com/)). 

Throughout the document, we report the $p$-value for several statistical tests. This does nothing more than provide a measure on the amount of evidence we have to suggest statistical significance. The closer the $p$-value is to 0 the stronger the evidence that our measure (correlation most of the time) is different than 0. 

**This paragraph would summarize some of the results for a quick reference**

## Analysis

### Income

```{r}
louis_income <- get_acs(geography = "zcta",
                        table = "B19101") %>% 
  filter(GEOID %in% zip_code_louis$zip,
         !variable %in% c("B19101_001"))

louis_income <- left_join(louis_income, geometry_zip_louis, by = "GEOID")
louis_income <- st_as_sf(louis_income, sf_column_name = "geometry")

louis_income <- left_join(louis_income, variables_2018[, 1:2], by = "variable")

louis_income$label <- as_factor(str_replace(louis_income$label, ".*!!(.*)", "\\1"))

louis_income_six_figure <- louis_income %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(GEOID, prop_100K) %>% 
  distinct(GEOID, prop_100K)

lvls_income <- louis_income_six_figure %>% 
  arrange(prop_100K) %>% 
  pull(GEOID)

louis_income <- left_join(louis_income, louis_covid, by = c("GEOID" = "zip"))
```


We look at the income data in increments designated by the Census Bureau. We calculated the proportion of people making over \$100,000 to order the zip codes on the $x$-axis in the following graph. 

```{r}
ggplot(louis_income) +
  geom_col(aes(factor(GEOID, levels = lvls_income), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Income Distribution for Each Louisville ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Income Level")
  
```

Zip code 40231 has no data, which is true throughout the analysis. We can compare the remaining zip codes with the case rate (per 100000) of COVID in the same zip codes. 

```{r}
pal <- colorNumeric(palette = "viridis", domain = louis_income$case_rate)
louis_income %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", louis_income$GEOID,
                            "</strong><br /> Case Rate ", signif(louis_income$case_rate, digits = 4)),
              weight = 4,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal(case_rate)) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~ case_rate,
            title = "Case Rate (per 100000)",
            opacity = 1)
```

```{r}
louis_income_cor <- left_join(louis_income_six_figure, louis_covid, by = c("GEOID"="zip"))
```

The next graph shows that it appears to have some correlation, i.e., the poorer the zip code the more likely to have a higher rate of infection. 

```{r}
ggplot(louis_income_cor, aes(prop_100K, case_rate)) +
  geom_point() +
  labs(title = "Slight Negative Correlation Between Infection and Income",
       subtitle = "ZIP code 40202 is an outlier.",
       x = "Proportion of Population that Makes >$100,000",
       y = "Case Rate (per 100,000)") +
  ggrepel::geom_label_repel(data = filter(louis_income_cor, case_rate > .03),
                            aes(label = GEOID))
```

```{r}
louis_income_cor_sans40202 <- louis_income_cor %>% 
  filter(case_rate < .03)
```

The 40202 ZIP code is a clear outlier. We remove this from the data to run a viable correlation test between the case rate and the proportion of the population that makes over \$100,000. We get a statistically significant ($p=$ `r cor.test(louis_income_cor_sans40202$prop_100K, louis_income_cor_sans40202$case_rate, use = "complete.obs")$p.value`) correlation value of `r cor.test(louis_income_cor_sans40202$prop_100K, louis_income_cor_sans40202$case_rate, use = "complete.obs")$estimate`. This verifies that there is indeed a moderately negative correlation with case rate and higher incomes.


### Gini Index

The Gini index is a measure of income inequality calculated by the Census Bureau using various different factors. The closer to 0 the Gini index, the more equal income is dispersed. A Gini index of 1 would indicate perfect inequality where a single group or individual collects all of the income. 

```{r}
louis_gini <- get_acs(geography = "zcta",
                      table = "B19083")
louis_gini <- left_join(louis_gini, geometry_zip_louis, by = "GEOID")
louis_gini <- st_as_sf(louis_gini, sf_column_name = "geometry")
louis_gini <- louis_gini %>% 
  filter(GEOID %in% zip_code_louis$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable")

louis_gini$label <- as_factor(str_replace(louis_gini$label, ".*!!(.*)", "\\1"))
```

Below, we map the Gini index onto each ZIP code to give an idea of the variation between ZIP codes.

```{r}
pal_gini <- colorNumeric(palette = "viridis", domain = louis_gini$estimate)

louis_gini %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", louis_gini$GEOID,
                            "</strong><br /> Gini Index: ", louis_gini$estimate),
              weight = 2,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal_gini(estimate)) %>% 
  addLegend("bottomright",
            pal = pal_gini,
            values = ~ estimate,
            title = "Gini Index",
            opacity = 1)
```

We also provide a scatterplot to visualize the connection between COVID case rate and the Gini Index. As we can see, ZIP code 40202 is an outlier and will be removed in future analysis. 

```{r}
louis_gini_cor <- left_join(louis_gini, louis_covid, by = c("GEOID"="zip"))
ggplot(louis_gini_cor, aes(estimate, case_rate)) +
  geom_point() +
  labs(title = "No Discernible Correlation Between Gini Index and COVID",
       x = "Gini Index",
       y = "Case Rate (per 100,000)") +
  ggrepel::geom_label_repel(data = filter(louis_gini_cor, case_rate > .03),
                            aes(label = GEOID))
louis_gini_cor_sans40202 <- filter(louis_gini_cor, case_rate < .03)
```

After removing 40202, we see that there is no real correlation between the Gini Index and the rate of infection for COVID. We get a correlation of `r cor.test(louis_gini_cor_sans40202$estimate, louis_gini_cor_sans40202$case_rate, use = "complete.obs")$estimate` ($p=$ `
`r cor.test(louis_gini_cor_sans40202$estimate, louis_gini_cor_sans40202$case_rate, use = "complete.obs")$p.value`), and thus indistinguishable from 0. 

### Health Insurance Coverage

We looked at the how much of the population has health insurance, regardless of whether it was public or private. The graph below shows the distribution of health insurance coverage across the Louisville ZIP codes. 

```{r}
louis_sex_age <- get_acs(geography = "zcta",
                         table = "B27001") %>% 
  filter(GEOID %in% zip_code_louis$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable")

louis_sex_age <- louis_sex_age %>% filter(str_count(label, "!!") >= 4)

louis_sex_age$label <- str_remove(louis_sex_age$label, "Estimate!!Total!!")

louis_sex_age <- separate(louis_sex_age,
                          label,
                          sep = "!!",
                          into = c("Sex", "Age", "HI_Coverage"))

louis_sex_age$HI_Coverage <- if_else(louis_sex_age$HI_Coverage == "No health insurance coverage", "No", "Yes")

louis_sex_age$Sex <- as_factor(louis_sex_age$Sex)
louis_sex_age$Age <- as_factor(louis_sex_age$Age)
louis_sex_age$HI_Coverage <- as_factor(louis_sex_age$HI_Coverage)  
```

```{r}
## Look at Health Insurance Coverage

louis_HI <- louis_sex_age %>% group_by(GEOID, HI_Coverage) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

### Graph Health Insurance Coverage by proportion

lvls_louis_HI <- louis_HI %>% 
  filter(HI_Coverage == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(louis_HI) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_HI), count, fill = HI_Coverage),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Health Care Coverage for Each Louisville ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Health Insurance \nCoverage?")
```

A vast majority of Louisville residents have health insurance, although there is quite a distinction between the 40218 ZIP code and the 40245 ZIP code. To see any correlation with COVID, we can look at a scatterplot.

```{r}
louis_HI_cor <- left_join(louis_HI, louis_covid, by = c("GEOID"="zip"))

filter(louis_HI_cor, HI_Coverage == "Yes") %>%
  ggplot(aes(prop, case_rate)) +
  geom_point() +
  labs(title = "Slight Correlation Between Health Insurance Coverage and COVID",
       x = "Proportion of Population that has Health Insurance",
       y = "Case Rate (per 100,000)") +
  ggrepel::geom_label_repel(data = filter(louis_HI_cor, case_rate > .03, HI_Coverage == "Yes"),
                            aes(label = GEOID))
```

```{r}
louis_HI_cor_sans40202 <- louis_HI_cor %>% 
  filter(HI_Coverage == "Yes",
         case_rate < .03)
```

There appears to be a definitive correlation, once we remove 40202. We end up with a moderate to strong correlation value of `r cor.test(louis_HI_cor_sans40202$prop, louis_HI_cor_sans40202$case_rate, use = "complete.obs")$estimate` ($p=$ `r signif(cor.test(louis_HI_cor_sans40202$prop, louis_HI_cor_sans40202$case_rate, use = "complete.obs")$p.value, digits = 2)`).

### Private Health Insurance vs. Public Health Insurance

Since access to private health insurance is not universal, we looked at the relationship between having some type of private health insurance and COVID as well as similarly looking at public health insurance. It is possible to have both public and private insurance, so these metrics are not mutually exclusive.

```{r}
louis_hi_private <- get_acs(geography = "zcta",
                            table = "B27002") %>% 
  filter(GEOID %in% zip_code_louis$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

louis_hi_public <- get_acs(geography = "zcta",
                           table = "B27003") %>% 
  filter(GEOID %in% zip_code_louis$zip) %>% 
  left_join(variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

louis_hi_private$label <- str_remove(louis_hi_private$label, "Estimate!!Total!!")
louis_hi_public$label <- str_remove(louis_hi_public$label, "Estimate!!Total!!")

louis_hi_private <- separate(louis_hi_private,
                             label,
                             sep = "!!",
                             into = c("Sex", "Age", "Private_HI"))
louis_hi_public <- separate(louis_hi_public,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Public_HI"))

louis_hi_private$Private_HI <- if_else(louis_hi_private$Private_HI == "No private health insurance", "No", "Yes")
louis_hi_public$Public_HI <- if_else(louis_hi_public$Public_HI == "No public coverage", "No", "Yes")

louis_hi_private$Sex <- as_factor(louis_hi_private$Sex)
louis_hi_public$Sex <- as_factor(louis_hi_public$Sex)
louis_hi_private$Age <- as_factor(louis_hi_private$Age)
louis_hi_public$Age <- as_factor(louis_hi_public$Age)
louis_hi_private$Private_HI <- as_factor(louis_hi_private$Private_HI)
louis_hi_public$Public_HI <- as_factor(louis_hi_public$Public_HI) 
```

#### Private Health Insurance

We breakdown each ZIP code by the proportion of residents who have private health insurance.

```{r}
louis_hi_private_2 <- louis_hi_private %>% group_by(GEOID, Private_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)

### Graph Private HI Coverage by proportion

lvls_louis_hi_private <- louis_hi_private_2 %>% 
  filter(Private_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(louis_hi_private_2) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_hi_private), count, fill = Private_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Prevalence of Private Health Insurance for Each Louisville ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Private Health \nInsurance Coverage?")
```

As we see, there is a wide range of those populations that have private health insurance, ranging from almost 90% of the population in 40245 to only 31.4% of the residents in 40210 having private health insurance. But how does this relate to COVID cases.

```{r}
louis_hi_private_cor <- left_join(louis_hi_private_2, louis_covid, by = c("GEOID"="zip")) %>% 
  filter(Private_HI == "Yes")

louis_hi_private_cor %>%
  ggplot(aes(prop, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_hi_private_cor, case_rate > .03),
                                          aes(label = GEOID)) +
  labs(title = "Possession of Private Health Insurance Seems \nto be Grouped by Another Factor",
       x = "Proportion of Population that has Private Health Insurance",
       y = "Case Rate (per 100,000)")
```

The graph shows that we don't have a correlation between the case rate of COVID and having private insurance. However, this grouping is interesting. It turns out it is geographically based as the next map shows. We split the proportion of private health insurance into three groups: low, medium, and high. The *low* group has less than 45% of its population possessing private health insurance. The *medium* group has between 45% and 65%, and the *high* group has over 65% of its population with private health insurance.

```{r}
louis_hi_private_cor <- louis_hi_private_cor %>% 
  mutate(hi_group = factor(if_else(prop < .5, "low",
                            if_else(prop < .65, "medium",
                                    "high")),
                           levels = c("low", "medium", "high")))
louis_hi_private_cor <- left_join(louis_hi_private_cor, geometry_zip_louis, by = "GEOID")
louis_hi_private_cor <- st_as_sf(louis_hi_private_cor, sf_column_name = "geometry")

pal_hi <- colorFactor(palette = c('purple', 'seagreen', 'red3'), domain = louis_hi_private_cor$hi_group)

louis_hi_private_cor %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", louis_hi_private_cor$GEOID,
                            "</strong><br /> Prop. w/ Private HI: ", signif(louis_hi_private_cor$prop, digits = 4)),
              weight = 4,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal_hi(hi_group)) %>% 
  addLegend("bottomright",
            pal = pal_hi,
            values = ~ hi_group,
            title = "Private Health \n Insurance Groups",
            opacity = 1)
```

#### Public Health Insurance

We breakdown each ZIP code by the proportion of residents who have public health insurance.

```{r}
louis_hi_public_2 <- louis_hi_public %>% group_by(GEOID, Public_HI) %>% 
  summarize(count = sum(estimate)) %>% 
  ungroup() %>% 
  group_by(GEOID) %>% 
  mutate(pop = sum(count),
         prop = count/pop)
### Graph Public HI Coverage by proportion

lvls_louis_hi_public <- louis_hi_public_2 %>% 
  filter(Public_HI == "Yes") %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(louis_hi_public_2) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_hi_public), count, fill = Public_HI),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Prevalence of Public Health Insurance for Each Louisville ZIP Code",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Public Health \nInsurance Coverage?")
```

We still have a descent distinction between ZIP codes as to the proportion of their populations that have public insurance, although it is not as pronounced (ranges from almost 10% to 72%). We can look at how these proportions relate to COVID in the following plot.

```{r}
louis_hi_public_cor <- left_join(louis_hi_public_2, louis_covid, by = c("GEOID"="zip"))

filter(louis_hi_public_cor, Public_HI == "Yes") %>%
  ggplot(aes(prop, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_hi_public_cor, 
                                          case_rate > .03,
                                          Public_HI == "Yes"),
                                          aes(label = GEOID)) +
  labs(title = "Public HI vs Case Rate")
```

We don't see the same groupings as we did before, but there are still 5 ZIP codes that are separated from the rest. Again, it looks as though geography (and all that it encompasses) explains this:

```{r}
louis_hi_public_cor <- louis_hi_public_cor %>% 
  mutate(hi_group = factor(if_else(prop < .5, "low", "high"),
                           levels = c("low", "high")))
louis_hi_public_cor <- left_join(louis_hi_public_cor, geometry_zip_louis, by = "GEOID")
louis_hi_public_cor <- st_as_sf(louis_hi_public_cor, sf_column_name = "geometry")

pal_hi <- colorFactor(palette = c('seagreen', 'red3'), domain = louis_hi_public_cor$hi_group)

louis_hi_public_cor %>% 
  st_transform(crs = "+init=epsg:4326") %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(provider = "CartoDB.Positron") %>% 
  addPolygons(popup = str_c("<strong>", louis_hi_public_cor$GEOID,
                            "</strong><br /> Prop. w/ Public HI: ", signif(louis_hi_public_cor$prop, digits = 4)),
              weight = 4,
              smoothFactor = 0,
              fillOpacity = 0.7,
              color = ~ pal_hi(hi_group)) %>% 
  addLegend("bottomright",
            pal = pal_hi,
            values = ~ hi_group,
            title = "Public Health \n Insurance Groups",
            opacity = 1)
```
### Age

We know people over 65 years of age are at higher risk of mortality from COVID. But we wanted to know if they are catching COVID at a higher rate. Is Louisville doing a good job of protecting their senior citizens? To answer this, we can look at the breakdown of ages in each of the ZIP codes and compare the proportion of senior citizens in each ZIP code to the case rate of COVID.

```{r}
louis_age <- louis_sex_age %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(Age %in% c("65 to 74 years",
                    "75 years and over")) %>% 
  mutate(prop_over65 = sum(prop)) %>% 
  arrange(prop_over65)
```

We start with our basic graphic displaying the distribution of ages throughout the ZIP codes.

```{r}
lvls_age <- louis_sex_age %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(Age %in% c("65 to 74 years",
                    "75 years and over")) %>% 
  mutate(prop_over65 = sum(prop)) %>% 
  arrange(prop_over65) %>% 
  select(GEOID, prop_over65) %>% 
  distinct(GEOID, prop_over65) %>% 
  pull(GEOID)
ggplot(louis_sex_age) +
  geom_col(aes(factor(GEOID, levels = lvls_age), estimate, fill = Age),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Age Distribution for Louisville ZIP Codes",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Age Ranges")
```

We focus on the relationship between COVID cases and the proportion of the population that is 65 years or older.

```{r}
louis_age_cor <- left_join(louis_age, louis_covid, by = c("GEOID"="zip"))

ggplot(louis_age_cor, aes(prop_over65, case_rate)) +
  geom_point() +
  labs(title = "Slight Correlation Between Proportion of Population Over 65 and COVID",
       x = "Proportion of Population Over Age 65",
       y = "Case Rate (per 100,000)") +
  ggrepel::geom_label_repel(data = filter(louis_age_cor, case_rate > .03)[1,],
                            aes(label = GEOID))
louis_age_cor_sans40202 <- filter(louis_age_cor, case_rate < .03)
```

There appears to be a slight correlation. After removal of the 40202 outlier, we get a correlation value of `r cor.test(louis_age_cor_sans40202$prop_over65, louis_age_cor_sans40202$case_rate, use = "complete.obs")$estimate` ($p=$ `r cor.test(louis_age_cor_sans40202$prop_over65, louis_age_cor_sans40202$case_rate, use = "complete.obs")$p.value`).

### Occupation

Many "essential" workers tend to be towards the lower end of the income bracket. We look at whether there is a relationship between essential workers and contracting COVID.

The Census Bureau breaks workers into five categories: Management, business, science and arts; service; sales and office; natural resources, construction, and maintenance; and production, transportation, and material moving. Although these categories undoubtedly may not be clear distinctions between essential and non-essential, we feel a good classification is to put natural resources, construction, and maintenance and production, transportation, material moving as essential and the other three categories as non-essential. 

```{r}
louis_occ <- get_acs(geography = "zcta",
                     table = "C24060") %>% 
  filter(GEOID %in% zip_code_louis$zip,
         variable %in% str_c("C24060_00", 2:6))
louis_occ <- left_join(louis_occ, variables_2018[, 1:2], by = "variable")
louis_occ$label <- as_factor(str_replace(louis_occ$label, ".*!!(.*)", "\\1"))
```

When we do, we get the following breakdown by ZIP code.

```{r}
louis_occ_ess <- louis_occ %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(str_detect(label,
                    "(Natural.*)|(Production.*)")) %>% 
  mutate(prop_ess = sum(prop)) %>% 
  select(GEOID, prop_ess) %>% 
  distinct(GEOID, prop_ess)

lvls_louis_ess <- louis_occ_ess %>% 
  arrange(prop_ess) %>% 
  pull(GEOID)

ggplot(louis_occ) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_ess), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupation Distribution for Louisville ZIP Codes",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Occupation")
```

ZIP code 40209 clearly has a distinguishing percentage of production, transportation, and material moving occupations. Unfortunately, Jefferson County Health Department has not COVID information on 40209. However, we still have a good spread of different occupations from ZIP code to ZIP code. We can compare the proportion of a ZIP code that has a designated essential job to the COVID case rate.

```{r}
louis_occ_cor <- left_join(louis_occ_ess, louis_covid, by = c("GEOID"="zip"))
ggplot(louis_occ_cor, aes(prop_ess, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_occ_cor, GEOID == "40202"),
                                          aes(label = GEOID)) +
  labs(title = "Essential Worker Proportion Slightly Related to COVID Case Rate",
       x = "Proportion of Population Deemed to Work Essential Jobs",
       y = "Case Rate (per 100,000)")
louis_occ_cor_sans40202 <- filter(louis_occ_cor, case_rate < .03)
```

We find a moderate correlation of `r cor.test(louis_occ_cor_sans40202$prop_ess, louis_occ_cor_sans40202$case_rate, use = "complete.obs")$estimate` ($p=$ `r cor.test(louis_occ_cor_sans40202$prop_ess, louis_occ_cor_sans40202$case_rate, use = "complete.obs")$p.value`).


### Public Transportation

The use of public transportation is more prevalent among lower incomes, while at the same time public transportation poses a higher risk of contracting COVID. We look to see if this idea plays out in the data for Louisville.

```{r}
louis_trans <- get_acs(geography = "zcta",
                       table = "B08301") %>% 
  filter(GEOID %in% zip_code_louis$zip,
         variable %in% c("B08301_002", str_c("B08301_0", c(10, 16:21))))

louis_trans <- left_join(louis_trans, variables_2018[, 1:2], by = "variable")
louis_trans$label <- as_factor(str_replace(louis_trans$label, ".*!!(.*)", "\\1"))

louis_trans <- louis_trans %>% 
  mutate(use_public = if_else(str_detect(label, "^Pub.*")|str_detect(label, "Taxicab"), TRUE, FALSE))

louis_trans <- left_join(louis_trans, louis_covid, by = c("GEOID"="zip"))
```

First, we look at the distribution of the population that uses public transportation broken down by ZIP.

```{r}
louis_trans_public <- louis_trans %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(use_public == TRUE) %>% 
  mutate(prop_public = sum(prop)) 

lvls_trans <- louis_trans_public %>% 
  arrange(prop_public) %>% 
  distinct(GEOID) %>% 
  pull(GEOID)

ggplot(louis_trans) +
  geom_col(aes(factor(GEOID, levels = lvls_trans), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Mode of Transporation for Louisville ZIP Codes",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Transporation")
```

For our analysis, we grouped "Public transportation (excluding taxicab)" and "Taxicab" together to form our public use variable. It's relationship to COVID can be seen in the following graph.

```{r}
louis_trans_public %>%
  filter(variable == "B08301_010") %>% 
  ggplot(aes(prop_public, case_rate)) +
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_trans_public, GEOID %in% c("40202", 
                                                                           "40203", 
                                                                           "40211",
                                                                           "40210"),
                                          variable == "B08301_010"),
                                          aes(label = GEOID)) +
  labs(title = "Use of Public Transportation Not Related to COVID Case Rate",
       x = "Proportion of Population Who Use Public Transportation",
       y = "Case Rate (per 100,000)")

louis_trans_public_sans40202 <- filter(louis_trans_public, case_rate < .03)
```

We notice a group of three ZIP codes with distinctly higher rates of public transportation use. These ZIP codes correspond to the downtown Louisville area. When we remove 40202 as an outlier and test for correlation, we find none (correlation value of `r cor.test(louis_trans_public_sans40202$prop_public, louis_trans_public_sans40202$case_rate, use = "complete.obs")$estimate` with $p=$ `r cor.test(louis_trans_public_sans40202$prop_public, louis_trans_public_sans40202$case_rate, use = "complete.obs")$p.value`).

### Citizenship

The Census Bureau provides a distinction between different types of U.S. citizenship (e.g., naturalization, born in U.S., born abroad, etc.). We focus on the simple distinction between being a U.S. citizen and not.

```{r}
louis_citizen_clean <- get_acs(geography = "zcta",
                         table = "B05001") %>% 
  filter(GEOID %in% zip_code_louis$zip,
         variable %in% str_c("B05001_00", 2:6))
  
louis_citizen <- left_join(louis_citizen_clean, variables_2018[, 1:2], by = "variable")
louis_citizen$label <- as_factor(str_replace(louis_citizen$label, ".*!!(.*)", "\\1"))

louis_citizen <- louis_citizen %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n)

louis_citizen <- left_join(louis_citizen, louis_covid, by = c("GEOID"="zip"))
```

Looking at the breakdown of the U.S. citizenship population in each ZIP code we see the expected high rates of citizenship in all ZIP codes. This ranges from 87.6% to 100% of the population being a U.S. citizen. 

```{r}
louis_citizen_prop_us <- louis_citizen %>% 
  filter(str_detect(label,
                    "^U.S.")) %>% 
  mutate(prop_us_citizen = sum(prop)) %>% 
  select(GEOID, prop_us_citizen) %>% 
  distinct(GEOID, prop_us_citizen)

lvls_louis_citizen <- louis_citizen_prop_us %>% 
  arrange(prop_us_citizen) %>% 
  pull(GEOID)

ggplot(louis_citizen) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_citizen), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "U.S. Citizenship for Louisville ZIP Codes",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Citizenship Status")
```

Even with high rates of citizenship, we see some correlation and grouping.

```{r}
louis_citizen <- left_join(louis_citizen, louis_citizen_prop_us, by = "GEOID")

ggplot(louis_citizen, aes(prop_us_citizen, case_rate)) + 
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_citizen, GEOID %in% c("40202",
                                                                      "40218",
                                                                      "40214",
                                                                      "40219"),
                                          variable == "B05001_002"), 
                            aes(label = GEOID)) +
  labs(title = "U.S. Citizenship Related to COVID Case Rate",
       x = "Proportion of Population U.S. Citizen",
       y = "Case Rate (per 100,000)")
louis_citizen_sans40202 <- filter(louis_citizen, case_rate < .03)
```

There are three ZIP codes that have a distinctively different proportion of their population which are U.S. citizens. These ZIP codes are connected geographically, but not in a seemingly meaningful way. This group, however, fits in with the trend line of the relationship by proportion of U.S. citizens and COVID cases. We get a moderate to strong correlation of `r cor.test(louis_citizen_sans40202$prop_us_citizen, louis_citizen_sans40202$case_rate, use = "complete.obs")$estimate` ($p=$ `r cor.test(louis_citizen_sans40202$prop_us_citizen, louis_citizen_sans40202$case_rate, use = "complete.obs")$p.value`)

### Race

Lastly, we looked at racial breakdown in each ZIP code. We've heard reports about black people being hit harder by COVID. We look to see if the data supports that in Louisville.

```{r}
louis_race_clean <- get_acs(geography = "zcta",
                            table = "B02001") %>% 
  filter(GEOID %in% zip_code_louis$zip,
         variable %in% str_c("B02001_00", 2:8))

louis_race <- left_join(louis_race_clean, variables_2018[, 1:2], by = "variable")
louis_race$label <- as_factor(str_replace(louis_race$label, ".*!!(.*)", "\\1"))

louis_race <- louis_race %>% 
  group_by(GEOID) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n)

louis_race <- left_join(louis_race, louis_covid, by = c("GEOID"="zip"))
```

We view the racial breakdown in each ZIP code. There are significant differences in several of the ZIP codes.

```{r}
lvls_louis_race <- louis_race %>% 
  filter(str_detect(label, "White alone")) %>% 
  arrange(prop) %>% 
  pull(GEOID)

ggplot(louis_race) +
  geom_col(aes(factor(GEOID, levels = lvls_louis_race), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Racial Makeup for Louisville ZIP Codes",
       x = "ZIP Code",
       y = "Proportion of Population", 
       fill = "Race")
```

For the most part, the rates of "White alone" and "Black or African American alone" are inversely related as the other races remain somewhat constant. As a result, we can focus on the "White alone" population as a measure of the racial breakdown for each ZIP and compare it to the COVID cases.

```{r}
louis_race %>% 
  filter(label == "White alone") %>% 
ggplot(aes(prop, case_rate)) + 
  geom_point() +
  ggrepel::geom_label_repel(data = filter(louis_race, prop < .25 | case_rate > .03,
                                          label == "White alone"), 
                            aes(label = GEOID)) +
  labs(title = "Racial Makeup of ZIP is Related to COVID Case Rate",
       x = "Proportion of Population Classified as 'White alone'",
       y = "Case Rate (per 100,000)")
louis_race_cor <- filter(louis_race, label == "White alone")
```

We see a moderate relationship between racial dispersion in ZIP codes and COVID with a correlation value of `r cor.test(louis_race_cor$prop, louis_race_cor$case_rate, use = "complete.obs")$estimate` ($p=$ `r cor.test(louis_race_cor$prop, louis_race_cor$case_rate, use = "complete.obs")$p.value`).


## Conclusions

Here we would put in our conclusions.
