---
title: "Milwaukee Metropolitan COVID-19 Data"
author: "Timmy Miller"
date: "9/27/2020"
output:
  html_document:
    df_print: paged
---


The COVID-19 pandemic disproportionately impacts lower-income communities. The Milwaukee Metropolitan area, consisting of Milwaukee, Washington, Waukesha, and Ozaukee counties, exemplifies this disparity. Milwaukee County, which has a much greater non-white population percentage and a lower median household income than the other metropolitan counties, expereinces more cases and deaths per 100000 people than the rest of the metropolitan counties. When taking into consideration conditions or demographics pre-exposed to death from COVID-19, smoking is much more prevalent in Milwaukee County, which could explain some of the difference in deaths between counties. However, the average age in Milwaukee County is much lower than the other counties, which is surprising given the rate at which individuals 65 and older die from COVID in comparison to the rest of society. Health insurance coverage rate is also slightly lower in Milwaukee County than in the rest of the Metropolitan area. Unfortunately, downloadable ZIP code data for Milwaukee was not able to be found at this time, so a more specific analysis could not be performed. 



```{r, echo=FALSE, message=FALSE}
#packages required
library(tidyverse)
library(readxl)
library(ggrepel)
library(gridExtra)
library(ggplot2)
library(dplyr)
library(tidycensus)
source("../Scripts/functions.R")
```





```{r, echo=FALSE}
#data set for Milwaukee Metropolitan COVID data, use the ../ to back up a folder
Milwaukee_Covid_Data <- read_excel("../Data/Milwaukee Covid Data.xls")


```




```{r, echo=FALSE}

CASES <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=CASES_PER_HUNDTHOU))+geom_col(fill="blue")+xlab("County")+ylab("Cases Per 100000 People")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Case Rates")

DEATHS <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=DEATHS_HUNDTHOU))+geom_col(fill="red")+xlab("County")+ylab("Deaths Per 100000 People")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Death Rates")

WHITEPOP <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=WHITEPOP))+geom_col(fill="orange")+xlab("County")+ylab("% of Population that is White")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="White Population Proportion")

MHI <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=MHI))+geom_col(fill="green")+xlab("County")+ylab("Median Household Income ($)")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Median Household Income")

SMOKE<-ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=PCT_SMK))+geom_col(fill="black")+xlab("County")+ylab("% of Population that Smokes")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Smoking Proportion")

OLDER_PPL<-ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=PCT_POP_65))+geom_col(fill="yellow")+xlab("County")+ylab("% of Population 65 or Older")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Age Information")

INSURANCE <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=Pct18to64HI))+geom_col(fill="purple")+xlab("County")+ylab("% of Population with Health Insurance")+theme(axis.text.x = element_text(angle = 45, hjust = 1))+labs(title="Health Insurance Data")
```








GINI Index Data
```{r, echo=FALSE}
milwaukee_gini <- get_acs(geography = "county", table = "B19083", state= "WI")
milwaukee_gini<- filter(milwaukee_gini, NAME=="Milwaukee County, Wisconsin" | NAME=="Ozaukee County, Wisconsin" | NAME=="Waukesha County, Wisconsin" | NAME=="Washington County, Wisconsin")
Milwaukee_Covid_Data <- merge(Milwaukee_Covid_Data, milwaukee_gini, by = "NAME")


GINI<-ggplot(milwaukee_gini)+geom_col(aes(x=NAME, y=estimate))+labs(title="GINI Index Data for Milwaukee Metropolitan Counties", x="County", y="GINI Index")+theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

Income Data
```{r, echo=FALSE}
milwaukee_income<- get_acs(geography="County", state="WI", table="B19101")
income <- load_variables(2018, "acs5")
income <- income[c(11995:12011),]
income<-rename(income, c("variable"="name"))
income <- merge(income, milwaukee_income, by="variable")
income <- filter( income, NAME=="Milwaukee County, Wisconsin" | NAME=="Ozaukee County, Wisconsin" | NAME=="Waukesha County, Wisconsin" | NAME=="Washington County, Wisconsin")
income$label <- as_factor(str_replace(income$label, ".*!!(.*)", "\\1"))
income <- income[c(5:68),]
```


```{r, echo=FALSE}
income_sixfig <- income %>% 
  group_by(NAME) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("$200,000 or more",
                      "$150,000 to $199,999",
                      "$125,000 to $149,999",
                      "$100,000 to $124,999")) %>% 
  mutate(prop_100K = sum(prop)) %>% 
  select(NAME, prop_100K) %>% 
  distinct(NAME, prop_100K)

lvls <-income_sixfig %>% 
  arrange(prop_100K) %>% 
  pull(NAME)

INCOME<-ggplot(income) +
  geom_col(aes(factor(NAME, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Median Household Income for Each \nMilwaukee Metropolitan County",
       x = "County",
       y = "Proportion", 
       fill = "Income Status")



```

Transportation Data
```{r, echo=FALSE}
wis_transport <- get_acs(geography="County", state="WI", table="B08301") %>%
  filter(NAME%in%c("Milwaukee County, Wisconsin", "Ozaukee County, Wisconsin", "Waukesha County, Wisconsin", "Washington County, Wisconsin"), 
        variable %in% c("B08301_002", str_c("B08301_0", c(10, 16:21))))

mil_transport <- left_join(wis_transport, variables_2018[,1:2], by="variable")
mil_transport$label <- as_factor(str_replace(mil_transport$label, ".*!!(.*)", "\\1"))
  
  
  



```

```{r, echo=FALSE}
transport_mil <- mil_transport %>% 
  group_by(NAME) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Public transportation (excluding taxicab)",
                      "Taxicab")) %>% 
  mutate(prop_pub = sum(prop)) %>% 
  select(NAME, prop_pub) %>% 
  distinct(NAME, prop_pub)

lvls <- transport_mil %>% 
  arrange(prop_pub) %>% 
  pull(NAME)
#lvls needs to be a vector (1 dimension) not a data frame (2 dimensions)

TRANSPORT<-ggplot(mil_transport) +
  geom_col(aes(factor(NAME, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Transportation Distribution for Each \nMilwaukee Metropolitan County",
       x = "County",
       y = "Proportion", 
       fill = "Transportation")
```


Occupation
```{r, echo=FALSE}
milwaukee_occupation <- get_acs(geography="County", state="WI", table="C24060")
occupation <- load_variables(2018, "acs5")
occupation <- occupation[c(26451:26486),]
occupation<-rename(occupation, c("variable"="name"))
occupation<- merge(occupation, milwaukee_occupation, by="variable")
occupation <- filter( occupation, NAME=="Milwaukee County, Wisconsin" | NAME=="Ozaukee County, Wisconsin" | NAME=="Waukesha County, Wisconsin" | NAME=="Washington County, Wisconsin")
occupation<- occupation[c(5:144),]
occupation$label <- as_factor(str_replace(occupation$label, ".*!!(.*)", "\\1"))

```


```{r, echo=FALSE}
wis_occupation <- get_acs(geography="County", state="WI", table="C24060") %>%
  filter(NAME%in%c("Milwaukee County, Wisconsin", "Ozaukee County, Wisconsin", "Waukesha County, Wisconsin", "Washington County, Wisconsin"), 
        variable %in% c("C24060_002","C24060_003", "C24060_004", "C24060_005","C24060_006"))

mil_occupation <- left_join(wis_occupation, variables_2018[,1:2], by="variable")
mil_occupation$label <- as_factor(str_replace(mil_occupation$label, ".*!!(.*)", "\\1"))
```



```{r, echo=FALSE}
occupation <- mil_occupation %>% 
  group_by(NAME) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Natural resources, construction, and maintenance occupations",
                      "Production, transportation, and material moving occupations")) %>% 
  mutate(prop_essen = sum(prop)) %>% 
  select(NAME, prop_essen) %>% 
  distinct(NAME, prop_essen)

lvls <- occupation %>% 
  arrange(prop_essen) %>% 
  pull(NAME)

OCCUPATION<-ggplot(mil_occupation) +
  geom_col(aes(factor(NAME, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Occupations for Each Milwaukee Metropolitan County",
       x = "County",
       y = "Proportion", 
       fill = "Occupation")
```

Citizenship
```{r, echo=FALSE}
milwaukee_citizenship <- get_acs(geography="County", state="WI", table="B05001")
citizenship <- load_variables(2018, "acs5")
citizenship <- citizenship[c(946:957),]
citizenship<-rename(citizenship, c("variable"="name"))
citizenship<- merge(citizenship, milwaukee_citizenship, by="variable")
citizenship <- filter( citizenship, NAME=="Milwaukee County, Wisconsin" | NAME=="Ozaukee County, Wisconsin" | NAME=="Waukesha County, Wisconsin" | NAME=="Washington County, Wisconsin")
citizenship$label <- as_factor(str_replace(citizenship$label, ".*!!(.*)", "\\1"))
citizenship<-citizenship[c(5:24),]



```

```{r, echo=FALSE}
citi_status <- citizenship %>% 
  group_by(NAME) %>% 
  mutate(n = sum(estimate),
         prop = estimate/n) %>% 
  filter(label %in% c("Not a U.S. citizen")) %>% 
  mutate(prop_not_citizen = sum(prop)) %>% 
  select(NAME, prop_not_citizen) %>% 
  distinct(NAME, prop_not_citizen)

lvls <- citi_status %>% 
  arrange(prop_not_citizen) %>% 
  pull(NAME)



ggplot(citizenship) +
  geom_col(aes(factor(NAME, levels = lvls), estimate, fill = label),
           position = "fill") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1)) +
  labs(title = "Citizenship Status for Each Milwaukee Metropolitan County",
       x = "County",
       y = "Proportion", 
       fill = "Citizenship Status")
```


```{r, echo=FALSE, fig.width=18, fig.height=25}
#now you can arrange these five charts in a combined graphic for publication
grid.arrange(CASES, DEATHS, WHITEPOP, MHI, OLDER_PPL, SMOKE, INSURANCE, GINI, OCCUPATION, INCOME, TRANSPORT, top= "Milwaukee Metroplitan Area County COVID-19 Demographic Information", ncol=2)
```
