---
title: "Milwaukee Metropolitan COVID-19 Data"
author: "Timmy Miller"
date: "9/8/2020"
output: word_document
---


The COVID-19 pandemic disproportionately impacts lower-income communities. The Milwaukee Metropolitan area, consisting of Milwaukee, Washington, Waukesha, and Ozaukee counties, exemplifies this disparity. Milwaukee County, which has a much greater non-white population percentage and a lower median household income than the other metropolitan counties, expereinces more cases and deaths per 100000 people than the rest of the metropolitan counties. When taking into consideration conditions or demographics pre-exposed to death from COVID-19, smoking is much more prevalent in Milwaukee County, which could explain some of the difference in deaths between counties. However, the average age in Milwaukee County is much lower than the other counties, which is surprising given the rate at which individuals 65 and older die from COVID in comparison to the rest of society. Health insurance coverage rate is also slightly lower in Milwaukee County than in the rest of the Metropolitan area. Unfortunately, downloadable ZIP code data for Milwaukee was not able to be found at this time, so a more specific analysis could not be performed. 



```{r, echo=FALSE, message=FALSE}
#packages required
library(tidyverse)
library(readxl)
library(ggrepel)
library(gridExtra)
library(ggplot2)
```





```{r, echo=FALSE}
#data set for Milwaukee Metropolitan COVID data, use the ../ to back up a folder

Milwaukee_Covid_Data <- read_excel("../Data/Milwaukee Covid Data.xls")
```




```{r, echo=FALSE}
#Script for five plots comparing Cases/100000 people, Deaths/100000 people, Median Household Income, % of the Population who smokes, and % of the population older than 65 in the four counties of Metroplitan Milwaukee

CASES <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=CASES_PER_HUNDTHOU))+geom_col(fill="blue")+xlab("County")+ylab("Cases Per 100000 People")

DEATHS <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=DEATHS_HUNDTHOU))+geom_col(fill="red")+xlab("County")+ylab("Deaths Per 100000 People")

WHITEPOP <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=WHITEPOP))+geom_col(fill="orange")+xlab("County")+ylab("% of Population that is White")

MHI <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=MHI))+geom_col(fill="green")+xlab("County")+ylab("Median Household Income ($)")

SMOKE<-ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=PCT_SMK))+geom_col(fill="black")+xlab("County")+ylab("% of Population that Smokes")

OLDER_PPL<-ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=PCT_POP_65))+geom_col(fill="yellow")+xlab("County")+ylab("% of Population 65 or Older")

INSURANCE <- ggplot(data=Milwaukee_Covid_Data, aes(x=NAME, y=PctHI))+geom_col(fill="purple")+xlab("County")+ylab("% of Population with Health Insurance")
```



```{r, echo=FALSE, fig.width=12, fig.height=8}
#now you can arrange these five charts in a combined graphic for publication
grid.arrange(CASES, DEATHS, WHITEPOP, MHI, OLDER_PPL, SMOKE, INSURANCE, top= "Milwaukee Metroplitan Area County COVID-19 Demographic Information")
```








