---
title: "COVID Impact on Chicago Service Workers-EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("functions.R")
library(tidycensus)
library(tidyverse)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_chicago <- zip_code %>% filter(primary_city == "Chicago",
                                        state == "IL")
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
state_of_interest <- ("IL")
```

## Maya Frederick

Sex by age and citizenship status
```{r}
chicago_citizen <- get_acs(geography = "zcta",
                           year = 2018,
                           table = "B05003") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B05003_001"))
chicago_citizen <- left_join(chicago_citizen, variables_2018[, 1:2], by = "variable")

chicago_citizen <- chicago_citizen %>% filter(str_count(label, "!!") >= 4)
chicago_citizen$label <- str_remove(chicago_citizen$label, "Estimate!!Total!!") 

chicago_citizen <- chicago_citizen %>% filter(!str_detect(label,"Foreign born$"))
chicago_citizen$label <- str_remove(chicago_citizen$label, "Foreign born!!") 


chicago_citizen <- separate(chicago_citizen,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age", "Citizen"))

chicago_citizen$Citizen <- if_else(
  chicago_citizen$Citizen == "Not a U.S. citizen", "No", "Yes")
```

The following code will calculate the proportion of the ZIP code populations that fit into each category. We always need to group by GEOID since we want these numbers per ZIP code. Next, we calculate the total ZIP code population. After that, we do a series of `group_by()`, `mutate()` commands to be able to calculate the proportions for each categorical breakdown. Note the naming conventions for the new columns. Also note that I've created a new data frame to house these proportions as to not mess up the original data pulled from the ACS. You should do this for all possible one-way, two-way, three-way, etc. factorial combinations of the categories (e.g., Sex, Age, Citizen, Sex-Age, Sex-Citizen, Age-Citizen, Sex-Age-Citizen). Also note the algorithmic nature of constructing the factorial combinations. 

```{r}
chicago_citizen_2 <- chicago_citizen %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Sex) %>% 
  mutate(prop_sex = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age) %>% 
  mutate(prop_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Citizen) %>% 
  mutate(prop_cit = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age) %>% 
  mutate(prop_sex_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Citizen) %>% 
  mutate(prop_sex_cit = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age, Citizen) %>% 
  mutate(prop_age_cit = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age, Citizen) %>% 
  mutate(prop_sex_age_cit = sum(estimate)/zip_pop) %>% 
  ungroup()
```

```{r}
chicago_citizen_ex <- chicago_citizen_2 %>%
  filter(variable == "B05003_020") # female, over 18, citizen

chicago_citizen_ex <- chicago_citizen_ex %>%
  rename(prop_female = prop_sex,
         prop_over_18 = prop_age,
         prop_citizen = prop_cit,
         prop_female_over_18 = prop_sex_age,
         prop_female_citizen = prop_sex_cit,
         prop_over_18_citizen = prop_age_cit,
         prop_female_over_18_citizen = prop_sex_age_cit)
```


## Ethan Morlock
```{r}
# Sex by age (Hispanic or Latino)
chicago_sex_age <- get_acs(geography = "zcta",
                           year = 2018,
                           table = "B01001I") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         !variable %in% c("B01001I_001", "B01001I_002", "B01001I_017"))
chicago_sex_age <- left_join(chicago_sex_age, variables_2018[, 1:2], by = "variable")
chicago_sex_age$label <- str_remove(chicago_sex_age$label, "Estimate!!Total!!")
chicago_sex_age <- separate(chicago_sex_age,
                            label,
                            sep = "!!",
                            into = c("Sex", "Age"))
chicago_sex_age$Age <- if_else(chicago_sex_age$Age %in% c('Under 5 years', '5 to 9 years', '10 to 14 years', '15 to 17 years'), "0-17 years",
                               if_else(chicago_sex_age$Age %in% c('18 and 19 years', '20 to 24 years', '25 to 29 years', '30 to 34 years'), "18-34 years",
                                       if_else(chicago_sex_age$Age %in% c('35 to 44 years', '45 to 54 years', '55 to 64 years'), "35-64 years", "65+")))
chicago_sex_age$Sex <- as_factor(chicago_sex_age$Sex)
chicago_sex_age$Age <- as_factor(chicago_sex_age$Age)

chicago_sex_age_2 <- chicago_sex_age %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Sex) %>% 
  mutate(prop_sex = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age) %>% 
  mutate(prop_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age) %>% 
  mutate(prop_sex_age = sum(estimate)/zip_pop) %>% 
  ungroup()

chicago_sex_age_ex <- chicago_sex_age_2 %>%
  filter(variable == "B01001I_022") # female, 18-34 

chicago_sex_age_ex <- chicago_sex_age_ex %>%
  rename(zip_pop_hisp = zip_pop,
         prop_female_hisp = prop_sex,
         prop_18_to_34_hisp = prop_age,
         prop_female_18_to_34_hisp = prop_sex_age)
```

```{r}
# Health insurance coverage status by sex by age 
chicago_HI_orig <- get_acs(geography = "zcta",
                           year = 2018,
                           table = "B27001") %>% 
  filter(GEOID %in% zip_code_chicago$zip)

chicago_HI <- left_join(chicago_HI_orig, variables_2018[, 1:2], by = "variable") %>% 
  filter(str_count(label, "!!") >= 4)

chicago_HI$label <- str_remove(chicago_HI$label, "Estimate!!Total!!")
chicago_HI <- separate(chicago_HI,
                       label,
                       sep = "!!",
                       into = c("Sex", "Age", "HI_coverage"))
chicago_HI$HI_coverage <- if_else(chicago_HI$HI_coverage == "With health insurance coverage", "Yes", "No")
chicago_HI$Age <- if_else(chicago_HI$Age %in% c('Under 6 years', '6 to 18 years'), "0-18 years",
                          if_else(chicago_HI$Age %in% c('19 to 25 years', '26 to 34 years'), "19-34 years",
                                  if_else(chicago_HI$Age %in% c('35 to 44 years', '45 to 54 years', '55 to 64 years'), "35-64 years", "65+")))
chicago_HI$Sex <- as_factor(chicago_HI$Sex)
chicago_HI$Age <- as_factor(chicago_HI$Age)
chicago_HI$HI_coverage <- as_factor(chicago_HI$HI_coverage)

chicago_HI_2 <- chicago_HI %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Sex) %>% 
  mutate(prop_sex = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age) %>% 
  mutate(prop_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, HI_coverage) %>% 
  mutate(prop_HI_cov = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age) %>% 
  mutate(prop_sex_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, HI_coverage) %>% 
  mutate(prop_sex_HI_cov = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age, HI_coverage) %>% 
  mutate(prop_age_HI_cov = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age, HI_coverage) %>% 
  mutate(prop_sex_age_HI_cov = sum(estimate)/zip_pop) %>% 
  ungroup()

chicago_HI_ex <- chicago_HI_2 %>%
  filter(variable == "B27001_038") # female, 19-34 year olds, yes

chicago_HI_ex <- chicago_HI_ex %>%
  rename(prop_female = prop_sex,
         prop_19_to_34 = prop_age,
         prop_with_HI = prop_HI_cov,
         prop_female_19_to_34 = prop_sex_age,
         prop_female_with_HI = prop_sex_HI_cov,
         prop_19_to_34_with_HI = prop_age_HI_cov,
         prop_female_19_to_34_with_HI = prop_sex_age_HI_cov)
```

```{r}
# Race (Hispanic or Latino)
chicago_race <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B03002") %>% 
  filter(GEOID %in% zip_code_chicago$zip,
         variable %in% c(str_c("B03002_00", 3:9), str_c("B03002_0", 13:19)))
chicago_race <- left_join(chicago_race, variables_2018[, 1:2], by = "variable")
chicago_race$label <- str_remove(chicago_race$label, "Estimate!!Total!!")
chicago_race <- separate(chicago_race,
                         label,
                         sep = "!!",
                         into = c("Hispanic_or_Latino", "Race"))
chicago_race$Hispanic_or_Latino <- if_else(chicago_race$Hispanic_or_Latino == "Not Hispanic or Latino", "No", "Yes")
chicago_race$Race <- if_else(chicago_race$Race == "White alone", "White",
                             if_else(chicago_race$Race == "Black or African American alone", "Black",
                                     if_else(chicago_race$Race == "American Indian and Alaska Native alone", "AIAN", 
                                             if_else(chicago_race$Race == "Asian alone", "Asian", 
                                                     if_else(chicago_race$Race == "Native Hawaiian and Other Pacific Islander alone", "NHPI", 
                                                             if_else(chicago_race$Race == "Some other race alone", "Other", "Two or more"))))))
chicago_race$Hispanic_or_Latino <- as_factor(chicago_race$Hispanic_or_Latino)
chicago_race$Race <- as_factor(chicago_race$Race)

chicago_race_2 <- chicago_race %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Hispanic_or_Latino) %>% 
  mutate(prop_His_Lat = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Race) %>% 
  mutate(prop_Race = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Hispanic_or_Latino, Race) %>% 
  mutate(prop_His_Lat_Race = sum(estimate)/zip_pop) %>% 
  ungroup()

chicago_race_ex <- chicago_race_2 %>%
  filter(variable == "B03002_003") # no, white

chicago_race_ex <- chicago_race_ex %>%
  rename(prop_not_His = prop_His_Lat,
         prop_White = prop_Race,
         prop_not_His_White = prop_His_Lat_Race)
```


## Pearl Muensterman
```{r}
#Sex by Industry
chicago_sex_by_ind_orig <- get_acs(geography = "zcta",
                                   year = 2018,
                                   table = "C24030") %>% 
  filter(GEOID %in% zip_code_chicago$zip)


chicago_sex_by_ind <- chicago_sex_by_ind_orig %>% 
  filter(variable %in% c(str_c("C24030_00", c(3, 6:9)), 
                         str_c("C24030_0", c(10, 13:14, 17, 21, 25:28, 30, 33:37, 40:41, 44, 48, 52:55))))

chicago_sex_by_ind <- left_join(chicago_sex_by_ind, variables_2018[, 1:2], by = "variable")

chicago_sex_by_ind$label <- str_remove(chicago_sex_by_ind$label, "Estimate!!Total!!")
chicago_sex_by_ind$label <- str_remove(chicago_sex_by_ind$label, "Arts, entertainment, and recreation, and accommodation and food services!!")

chicago_sex_by_ind <- separate(chicago_sex_by_ind,
                               label,
                               sep = "!!",
                               into = c("Sex", "Industry"))

chicago_sex_by_ind$Sex <- as.factor(chicago_sex_by_ind$Sex)
chicago_sex_by_ind$Industry <- as.factor(chicago_sex_by_ind$Industry)


chicago_sex_by_ind <- chicago_sex_by_ind %>% mutate(Service_Worker = chicago_sex_by_ind$Industry)

chicago_sex_by_ind$Service_Worker <- if_else((chicago_sex_by_ind$Service_Worker == "Construction" | chicago_sex_by_ind$Service_Worker == "Manufacturing" | chicago_sex_by_ind$Service_Worker == "Retail trade" | chicago_sex_by_ind$Service_Worker == "Transportation and warehousing, and utilities" | chicago_sex_by_ind$Service_Worker == "Accommodation and food services"), "Yes", "No")

chicago_sex_by_ind <- select(chicago_sex_by_ind, -Industry, -moe)

chicago_sex_by_ind_2 <- chicago_sex_by_ind %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate, na.rm = TRUE)) %>% 
  group_by(GEOID, Sex) %>% 
  mutate(prop_sex = sum(estimate, na.rm = TRUE)/zip_pop) %>% 
  group_by(GEOID, Service_Worker) %>% 
  mutate(prop_service = sum(estimate, na.rm = TRUE)/zip_pop) %>% 
  group_by(GEOID, Sex, Service_Worker) %>% 
  mutate(prop_sex_service = sum(estimate, na.rm = TRUE)/zip_pop) %>% 
  ungroup()

#Male, Service Worker
chicago_sex_by_ind_pick <- chicago_sex_by_ind_2 %>% 
  filter(variable == "C24030_006")

chicago_sex_by_ind_pick <- chicago_sex_by_ind_pick %>%
  rename(zip_pop_work_eligible = zip_pop,
         prop_male_work_eligible = prop_sex,
         prop_service = prop_service,
         prop_male_work_eligible_service =  prop_sex_service)
```


```{r}
#Sex by Age
chicago_sex_by_age <- get_acs(geography = "zcta",
                              year = 2018,
                              table = "B01001") %>% 
  filter(GEOID %in% zip_code_chicago$zip)

chicago_sex_by_age <- left_join(chicago_sex_by_age, variables_2018[, 1:2], by = "variable")
chicago_sex_by_age <- chicago_sex_by_age %>% filter(str_count(label, "!!") == 3)

chicago_sex_by_age$label <- str_remove(chicago_sex_by_age$label, "Estimate!!Total!!")

chicago_sex_by_age <- separate(chicago_sex_by_age,
                               label,
                               sep = "!!",
                               into = c("Sex", "Age"))

chicago_sex_by_age$Sex <- as.factor(chicago_sex_by_age$Sex)
chicago_sex_by_age$Age <- as.factor(chicago_sex_by_age$Age)

chicago_sex_by_age <- chicago_sex_by_age %>% 
  mutate(Age_Group = case_when(
    variable %in% c(str_c("B01001_00", 3:6), str_c("B01001_0", 27:30)) ~ "Under 5 to 17",
    variable %in% c(str_c("B01001_00", 7:9), str_c("B01001_0", 10:12), 
                    str_c("B01001_0", 31:36)) ~ "18 to 34",
    variable %in% c(str_c("B01001_0", 13:19), str_c("B01001_0", 37:43)) ~ "35 to 64",
    variable %in% c(str_c("B01001_0", 20:25), str_c("B01001_0", 44:49)) ~ "65 and Older"))

chicago_sex_by_age <- subset(chicago_sex_by_age, select = -Age)

chicago_sex_by_age_2 <- chicago_sex_by_age %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Sex) %>% 
  mutate(prop_sex = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age_Group) %>% 
  mutate(prop_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Sex, Age_Group) %>% 
  mutate(prop_sex_age = sum(estimate)/zip_pop) %>% 
  ungroup()

#Female, 18-34 years
chicago_sex_by_age_pick <- chicago_sex_by_age_2 %>% 
  filter(variable == "B01001_031")

chicago_sex_by_age_pick <- chicago_sex_by_age_pick %>% 
  rename(prop_female = prop_sex,
         prop_18_to_34 = prop_age,
         prop_female_18_to_34 = prop_sex_age)
```


## Timmy Miller

```{r}
#ancestry


ancestry<-left_join((get_acs(geography="zcta", table="B04006")), variables_2018, by="variable")

ancestry$label<- as_factor(str_replace(ancestry$label, ".*!!(.*)", "\\1"))

ancestry<-filter(ancestry, label!="Total")

ancestry<- filter(ancestry, GEOID %in% 60601:60827)



ancestry_2 <- ancestry %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, label) %>% 
  mutate(prop_anc = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, label)%>% 
  ungroup()

#The next line doesn't run since chicago_covid doesn't exist. 
#ancestry<- left_join(ancestry, chicago_covid, by=c("GEOID"="Zip"))

```

```{r}
#health insurance by type 

HI_type_orig<-left_join(
  (get_acs(geography="zcta", table="B27010")), variables_2018, by="variable") %>% 
  filter(GEOID %in% zip_code_chicago$zip)

HI_type_final<-HI_type_orig%>%filter(str_count(label, "!!") >=3, 
                                     !str_detect(label, "of health insurance coverage$"))
HI_type_final$label<-str_remove(HI_type_final$label, "Estimate!!Total!!")


HI_type_final <- separate(HI_type_final,
                          label,
                          sep = "!!",
                          into = c( "Age_Group", "Number_HI_polcies", "HI_Type"))

HI_type_final <- HI_type_final %>% 
  mutate(num_HI_policies = case_when(
    str_detect(Number_HI_polcies, "With one") ~ "One",
    str_detect(Number_HI_polcies, "With two") ~ "Two or more",
    TRUE ~ "None"
  )) %>% 
  select(-Number_HI_polcies, -concept)

HI_type_final$HI_Type <- replace_na(HI_type_final$HI_Type, "None")

HI_type_final_2 <- HI_type_final  %>% 
  arrange(GEOID) %>% 
  group_by(GEOID) %>% 
  mutate(zip_pop = sum(estimate)) %>% 
  group_by(GEOID, Age_Group) %>% 
  mutate(prop_age = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, num_HI_policies) %>% 
  mutate(prop_num_policies = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, HI_Type) %>% 
  mutate(prop_HI_type = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age_Group, num_HI_policies) %>% 
  mutate(prop_age_num_policies = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age_Group, HI_Type) %>% 
  mutate(prop_age_HI_type = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, num_HI_policies, HI_Type) %>% 
  mutate(prop_num_policies_HI_type = sum(estimate)/zip_pop) %>% 
  group_by(GEOID, Age_Group, num_HI_policies, HI_Type) %>% 
  mutate(prop_age_num_policies_HI_type = sum(estimate)/zip_pop) %>% 
  ungroup()

```

We now select the appropriate rows to create a tidy dataset to work with. 

```{r}
#19-34 years, one policy, employer-based and
#19-34 years, no policy, none
chicago_HI_type_ex <- HI_type_final_2 %>% 
  filter(variable %in% c("B27010_020", "B27010_023", "B27010_033"))

#Since we are selecting two rows per ZIP code, we need to move one of the proportions to its own column
# to create a tidy dataset
chicago_HI_type_ex <- pivot_wider(chicago_HI_type_ex, 
                                  names_from = variable, 
                                  values_from = c(prop_num_policies,
                                                  prop_HI_type,
                                                  prop_age_num_policies,
                                                  prop_age_HI_type,
                                                  prop_num_policies_HI_type,
                                                  prop_age_num_policies_HI_type)) %>%
  select(GEOID, zip_pop, starts_with("prop")) %>% 
  group_by(GEOID) %>% 
  summarise_each(funs(first(.[!is.na(.)]))) #this collapses and consolidates the rows since 
# the pivot_wider() command introduced several NAs

chicago_HI_type_ex <- chicago_HI_type_ex %>% 
  rename_with(~ str_replace(., "B27010_020", "employer"), ends_with("B27010_020")) %>% 
  rename_with(~ str_replace(., "B27010_023", "medicaid"), ends_with("B27010_023")) %>%  
  rename_with(~ str_replace(., "B27010_033", "none"), ends_with("B27010_033"))

#This renames the columns appropriately and deletes duplicated columns
chicago_HI_type_ex <- chicago_HI_type_ex %>% 
  rename_with(~ str_replace(., "age", "19_to_34")) %>% 
  rename_with(~ str_replace(., "prop_HI_type", "prop")) %>% 
  rename_with(~ str_replace(., "num_policies_employer", "one_policy")) %>% 
  select(-contains("num_policies_medicaid"), 
         -prop_none,
         -prop_19_to_34_HI_type_none,
         -prop_num_policies_HI_type_none) %>% 
  rename_with(~ str_replace(., "HI_type_employer", "employer")) %>% 
  rename_with(~ str_replace(., "HI_type_medicaid", "medicaid")) %>% 
  rename(prop_no_HI = prop_num_policies_none,
         prop_19_to_34_no_HI = prop_19_to_34_num_policies_none) %>% 
  select(-prop_19_to_34_num_policies_medicaid,
         -prop_19_to_34_num_policies_employer,
         -starts_with("prop_num_policies_"),
         -prop_19_to_34_num_policies_HI_type_none) 
```


## Join all data frames together

Here, we join the final proportion columns chosen above to create a master data frame.

```{r}
chicago_master <- chicago_citizen_ex %>% 
  select(GEOID, zip_pop, starts_with("prop"))
chicago_master <- full_join(chicago_master, chicago_sex_age_ex, by = "GEOID")
chicago_master <- full_join(chicago_master, chicago_HI_ex, by = "GEOID")
chicago_master <- full_join(chicago_master, chicago_race_ex, by = "GEOID")
chicago_master <- full_join(chicago_master, chicago_sex_by_ind_pick, by = "GEOID")
chicago_master <- full_join(chicago_master, chicago_sex_by_age_pick, by = "GEOID")
chicago_master <- full_join(chicago_master, chicago_HI_type_ex, by = "GEOID")
chicago_master <- chicago_master %>% 
  select(GEOID, zip_pop.x, starts_with("zip_pop_"), starts_with("prop"), -prop_female.x, -prop_female.y, -prop_19_to_34.y, -prop_no_HI) %>% 
  rename_with(~ str_remove(., "\\.x"))
write_csv(chicago_master, "../Data/chicago_master_for_service_model.csv")
```

