---
title: "COVID Impact on Chicago Service Workers-EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("functions.R")
library(tidycensus)
library(tidyverse)
zip_code <- read_csv("../Data/zip_county.csv")
zip_code_chicago <- zip_code %>% filter(primary_city == "Chicago",
                                        state == "IL")
variables_2018 <- load_variables(2018, "acs5", cache = TRUE) %>%
  rename(variable = name)
state_of_interest <- ("IL")
```

## Maya Frederick



## Ethan Morlock
```{r}
# Sex by age (Hispanic or Latino)
chicago_sex_age <- get_acs(geography = "zcta",
                                   year = 2018,
                                   table = "B01001I") %>% 
                              filter(GEOID %in% zip_code_chicago$zip,
                                     !variable %in% c("B01001I_001", "B01001I_002", "B01001I_017"))
chicago_sex_age <- left_join(chicago_sex_age, variables_2018[, 1:2], by = "variable")
chicago_sex_age$label <- str_remove(chicago_sex_age$label, "Estimate!!Total!!")
chicago_sex_age <- separate(chicago_sex_age,
                    label,
                    sep = "!!",
                    into = c("Sex", "Age"))
chicago_sex_age$Sex <- as_factor(chicago_sex_age$Sex)
chicago_sex_age$Age <- as_factor(chicago_sex_age$Age)
write_csv(chicago_sex_age, "../Data/chicago_sex_age.csv")

# Health insurance coverage status by sex by age 
chicago_HI <- get_acs(geography = "zcta",
                   year = 2018,
                   table = "B27001") %>% 
              filter(GEOID %in% zip_code_chicago$zip,
                     !variable %in% c("B27001_001", "B27001_002", "B27001_003", "B27001_006", "B27001_009", "B27001_012",
                                      "B27001_015", "B27001_018", "B27001_021", "B27001_024", "B27001_027", "B27001_030", "B27001_031",
                                      "B27001_034", "B27001_037", "B27001_040", "B27001_043", "B27001_046", "B27001_049", "B27001_052",
                                      "B27001_055"))
chicago_HI <- left_join(chicago_HI, variables_2018[, 1:2], by = "variable")
chicago_HI$label <- str_remove(chicago_HI$label, "Estimate!!Total!!")
chicago_HI <- separate(chicago_HI,
                    label,
                    sep = "!!",
                    into = c("Sex", "Age", "HI_coverage"))
chicago_HI$HI_coverage <- if_else(chicago_HI$HI_coverage == "With health insurance coverage", "Yes", "No")
chicago_HI$Sex <- as_factor(chicago_HI$Sex)
chicago_HI$Age <- as_factor(chicago_HI$Age)
chicago_HI$HI_coverage <- as_factor(chicago_HI$HI_coverage)
write_csv(chicago_HI, "../Data/chicago_HI.csv")

# Race (Hispanic or Latino)
chicago_race <- get_acs(geography = "zcta",
                     year = 2018,
                     table = "B03002") %>% 
                filter(GEOID %in% zip_code_chicago$zip,
                       variable %in% c("B03002_010", "B03002_011", "B03002_020", "B03002_021"))
chicago_race <- left_join(chicago_race, variables_2018[, 1:2], by = "variable")
chicago_race$label <- str_remove(chicago_race$label, "Estimate!!Total!!")
chicago_race <- separate(chicago_race,
                    label,
                    sep = "!!",
                    into = c("Hispanic_or_Latino", "Two_or_more", "Including_other_race"))
chicago_race$Hispanic_or_Latino <- if_else(chicago_race$Hispanic_or_Latino == "Not Hispanic or Latino", "No", "Yes")
chicago_race$Including_other_race <- if_else(chicago_race$Including_other_race == "Two races including Some other race", "Including", "Excluding")
chicago_race$Hispanic_or_Latino <- as_factor(chicago_race$Hispanic_or_Latino)
chicago_race$Two_or_more <- as_factor(chicago_race$Two_or_more)
chicago_race$Including_other_race <- as_factor(chicago_race$Including_other_race)
write_csv(chicago_race, "../Data/chicago_race.csv")
```


## Pearl Muensterman
```{r}
#Occupation by Class Worker
chicago_occ_by_class <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "C24060") %>% 
                  filter(GEOID %in% zip_code_chicago$zip,
                         variable %in% c("C24060_007", "C24060_009", "C24060_011",
                                         "C24060_012", "C24060_031","C24060_033",
                                         "C24060_035", "C24060_036"))
chicago_occ_by_class <- left_join(chicago_occ_by_class, variables_2018[, 1:2], by = "variable")
chicago_occ_by_class$label <- as_factor(str_replace(chicago_occ_by_class$label, ".*!!(.*)", "\\1"))
```
Employee of private company workers: total (C24060_007); service occupation (C24060_009); natural resources, construction, and maintenance occupations (C24060_011); and production, transportation, and material moving occupations (C24060_012). 
Self-employed in own not incorporated business workers and unpaid family workers: total (C24060_031); service occupation (C24060_033); natural resources, construction, and maintenance occupations (C24060_035); and production, transportation, and material moving occupations (C24060_036).

```{r}
#Sex by Age Total
chicago_sex_by_age <- get_acs(geography = "zcta",
                        year = 2018,
                        table = "B01001") %>% 
                  filter(GEOID %in% zip_code_chicago$zip,
                         variable %in% c("B01001_001"))
chicago_sex_by_age <- left_join(chicago_sex_by_age, variables_2018[, 1:2], by = "variable")
chicago_sex_by_age$label <- as_factor(str_replace(chicago_sex_by_age$label, ".*!!(.*)", "\\1"))
```



## Timmy Miller


